\documentclass[11pt]{article}

\usepackage[margin=1in]{geometry}
\usepackage{amsmath,amssymb,amsthm,mathtools}
\usepackage{booktabs,array}
\usepackage{enumitem}
\usepackage{xcolor}
\usepackage{hyperref}
\hypersetup{colorlinks=true,linkcolor=blue,citecolor=blue,urlcolor=blue}

\DeclareRobustCommand{\ANCHOR}{\textnormal{\textsc{Anchor}}}
\newcommand{\Unif}{\operatorname{Unif}}
\newcommand{\code}[1]{\texttt{#1}}
\newcommand{\lb}{\operatorname{lb}}
\newcommand{\ub}{\operatorname{ub}}
\newcommand{\Mult}{\operatorname{Mult}}
\newcommand{\Can}{\operatorname{Can}}
\newcommand{\pos}{\operatorname{pos}}
\newcommand{\Real}{\mathbb{R}}
\newcommand{\Nat}{\mathbb{N}}
\newcommand{\Znodes}{\mathcal Z}
\newcommand{\Callback}{\ensuremath{\mathsf{Callback}}}
\newcommand{\Request}{\ensuremath{\mathsf{Request}}}
\newcommand{\Active}{\ensuremath{\mathsf{Active}}}
\newcommand{\Src}{\ensuremath{\mathsf{Src}}}
\newcommand{\Lat}{\ensuremath{\mathsf{Lat}}}
\newcommand{\Stop}{\ensuremath{\mathsf{Stop}}}
\newcommand{\true}{\ensuremath{\mathsf{true}}}
\newcommand{\false}{\ensuremath{\mathsf{false}}}
\newcommand{\Btree}{\ensuremath{\mathsf{Btree}}}
\newcommand{\Atree}{\ensuremath{\mathsf{Atree}}}
\newcommand{\sideA}{\ensuremath{\mathsf A}}
\newcommand{\sideB}{\ensuremath{\mathsf B}}

\newtheorem{theorem}{Theorem}[section]
\newtheorem{lemma}[theorem]{Lemma}
\newtheorem{proposition}[theorem]{Proposition}
\newtheorem{corollary}[theorem]{Corollary}
\theoremstyle{definition}
\newtheorem{definition}[theorem]{Definition}
\newtheorem{remark}[theorem]{Remark}

\newenvironment{algorithmbox}[1]{%
  \par\medskip\noindent\hrule\smallskip
  \noindent\textbf{#1}\par\smallskip
  \begin{enumerate}[leftmargin=1.7em,itemsep=2pt,topsep=2pt]
}{%
  \end{enumerate}
  \smallskip\hrule\medskip
}

\newenvironment{algorithmsteps}{%
  \begin{enumerate}[leftmargin=1.7em,itemsep=2pt,topsep=2pt]
}{%
  \end{enumerate}
}

\title{\ANCHOR: Exact Independent Sampling for Cross-Color Box-Intersection Joins}
\author{}
\date{}

\begin{document}
\maketitle

\begin{abstract}
\ANCHOR{} is an exact sampler for the cross-color intersection join of two sets of axis-aligned boxes in constant dimension.  A query of length $t$ returns an ordered sequence of $t$ independent uniform samples from the join, with replacement, or reports that the join is empty when a positive-length query has no valid output.  The algorithm decomposes the join by a current-dimensional anchored ownership rule, routes each anchor through a canonical dyadic cover of the opposite side, and recurses on suffix dimensions.  The final one-dimensional suffixes are represented by disjoint terminal atoms.  Exact multinomial quota allocation over pair-disjoint subproblems gives the target ordered i.i.d. distribution.  Three materialization levels are described: streaming, prefix caching, and full compilation.  In the exact-RAM model, after sorted views have been built, the streaming query time is $O(N\log^{d-1}(N+1)+t)$, PrefixCache$(k)$ uses $O(N\log^k(N+1))$ space with an adaptive suffix-query bound, and Compiled uses $O(N\log^{d-1}(N+1))$ space with $O(t)$ query time.
\end{abstract}

\section{Problem, conventions, and computation model}
\label{sec:setup}

\subsection{Input and join objective}

The input consists of two color classes of $d$-dimensional axis-aligned boxes,
\[
  \mathcal R \quad\text{and}\quad \mathcal S,
\]
where $d$ is a fixed constant.  A box $q$ is represented as a product of half-open intervals,
\[
  q=\prod_{j=1}^d [L_j(q),U_j(q)).
\]
In one dimension, two half-open intervals intersect if and only if
\[
  L_j(a)<U_j(b)
  \quad\text{and}\quad
  L_j(b)<U_j(a).
\]
Before any sorted view is constructed, every box $q$ with $L_j(q)\ge U_j(q)$ in at least one dimension $j$ is discarded.  Such a box has empty extent and cannot participate in a non-empty intersection.  All subsequent statements refer to the instance after this deterministic pruning step.  The object identities of the remaining boxes are immutable: geometrically identical boxes are still distinct input objects.

The cross-color intersection join is
\[
  J=\{(a,b)\in\mathcal R\times\mathcal S: a\sim b\},
\]
where $a\sim b$ means that $a$ and $b$ intersect in every dimension.  A query specifies a length
\[
  t\in\mathbb Z_{\ge0}
\]
and asks for an ordered sequence
\[
  \mathbf S=(S_1,\ldots,S_t).
\]
If $J\ne\emptyset$, the target distribution is
\[
  S_1,\ldots,S_t\overset{\mathrm{i.i.d.}}{\sim}\Unif(J),
\]
or, equivalently,
\[
  \Pr[\mathbf S=(p_1,\ldots,p_t)]=|J|^{-t}
  \quad\text{for every }(p_1,\ldots,p_t)\in J^t.
\]
Sampling is with replacement.  If $t=0$, the returned sequence is empty.  If $t>0$ and $J=\emptyset$, the returned value is the distinguished symbol \code{EMPTY-INSTANCE}.

\subsection{Suffix joins}

\ANCHOR{} processes dimensions in the order $1,2,\ldots,d$.  For two sets of boxes $A$ and $B$, define the suffix join
\[
  J^{(\ell)}(A,B)=
  \{(a,b)\in A\times B:
    a,b\text{ intersect in every dimension }\ell,\ell+1,\ldots,d\}.
\]
The top-level join is $J^{(1)}(\mathcal R,\mathcal S)$.  A recursive call at level $\ell$ enforces dimension $\ell$ by a pair-disjoint ownership decomposition and passes the remaining dimensions $\ell+1,\ldots,d$ to recursive calls.  The base level $\ell=d$ is handled directly by one-dimensional atoms.

A level-$\ell$ local instance is called \emph{prefix-certified} when every pair in its full cross product $A\times B$ already intersects in dimensions $1,2,\ldots,\ell-1$.  The top-level instance is prefix-certified vacuously.  The anchored local blocks constructed at level $\ell$ are prefix-certified for level $\ell+1$: the previous dimensions remain certified by induction, and the current dimension is certified by the ownership inequalities defining the local block.  Consequently, inside a level-$\ell$ recursive call, testing the suffix condition $J^{(\ell)}(A,B)$ is exactly the remaining work needed to identify full-dimensional join pairs within that local block.

\subsection{Sorted views and ranks}

A recursive call at level $\ell$ receives a sorted-view object
\[
  \mathfrak V_\ell(A,B)=
  (A^{L_\ell},A^{L_{\ell+1}},\ldots,A^{L_d},A^{U_d};
   B^{L_\ell},B^{L_{\ell+1}},\ldots,B^{L_d},B^{U_d}).
\]
Here $A^{L_j}$ is the list of objects in $A$ sorted by $L_j$, and $A^{U_d}$ is the list of objects in $A$ sorted by $U_d$; the $B$-side views are defined analogously.  Ties inside a sorted view are resolved by any fixed deterministic stable order.  The stable order among objects with equal left endpoint does not affect the set represented by any rank interval, because all interval boundaries are determined by strict or non-strict endpoint inequalities.

Only the last-dimensional upper-endpoint views $A^{U_d}$ and $B^{U_d}$ are carried in every recursive view.  For dimensions $\ell<d$, the current ownership rule and canonical routing need only the current left-endpoint ranks.  The endpoint $U_\ell(q)$ of an object $q$ is stored in the object record and used as a binary-search threshold in the opposite left-sorted view.  No sorted $U_\ell$ view is needed for $\ell<d$.  At the terminal level, the $U_d$-sorted views support monotone scans for the one-dimensional rank boundaries.

For a side $Y$, a dimension $\ell$, and an object $y\in Y$, let
\[
  \pos_Y^{(\ell)}(y)
\]
be the zero-based position of $y$ in $Y^{L_\ell}$.  For a threshold $\alpha$, define
\[
  \lb_Y^{(\ell)}(\alpha)=|\{y\in Y:L_\ell(y)<\alpha\}|,
  \qquad
  \ub_Y^{(\ell)}(\alpha)=|\{y\in Y:L_\ell(y)\le\alpha\}|.
\]
Thus $\lb$ is the first position whose left endpoint is at least $\alpha$, and $\ub$ is the first position whose left endpoint is strictly larger than $\alpha$.  When the dimension is clear, the superscript $(\ell)$ is omitted.

If the sorted views are not supplied, the top-level views can be built in
\[
  O(dN\log N)=O(N\log N)
\]
time for fixed $d$, where $N=|\mathcal R|+|\mathcal S|$.  All main bounds are stated after the top-level sorted views have been constructed; for raw unsorted input, the initial sorting cost is added separately.

\subsection{Exact discrete-random model}
\label{subsec:exact-model}

All guarantees are stated in an exact-RAM model.  Endpoint comparisons are exact.  Cardinalities, offsets, and weights are exact nonnegative integers.  A word operation on such values is counted as one RAM operation; bit-level costs are described separately at the end of this subsection.  The random choices made by distinct primitive invocations are mutually independent unless they are explicitly coupled by a common quota vector or a common shuffle.

The following exact discrete-random primitives are available.
\begin{description}[leftmargin=1.8em,itemsep=2pt,topsep=2pt]
\item[Uniform integer.] Given integers $p<q$, draw uniformly from $\{p,p+1,\ldots,q-1\}$ in $O(1)$ time.
\item[Weighted index.] Given $k$ nonnegative integer weights $w_1,\ldots,w_k$ with $W=\sum_iw_i>0$, \code{BuildIndex} constructs, in $O(k)$ time and $O(k)$ words, a sampler returning label $i$ with probability $w_i/W$ in $O(1)$ time.  Zero-weight labels are never returned.
\item[Multinomial quotas.] Given $h\ge0$ and weights $w_1,\ldots,w_k$ with total $W>0$, \code{DrawQuotas} returns
\[
  (H_1,\ldots,H_k)\sim
  \Mult\left(h;\frac{w_1}{W},\ldots,\frac{w_k}{W}\right)
\]
exactly.  It is implemented by one exact weighted-index construction, $h$ independent exact weighted draws, and exact counting of the returned labels.  The cost is $O(k+h)$ time and $O(k)$ space.
\item[Shuffle.] Fisher--Yates shuffling produces a uniformly random permutation of a finite sequence in linear time.
\end{description}
No floating-point probabilities are used.  The displayed $O(t)$ query bound for the fully compiled representation uses the exact $O(1)$ weighted-index primitive.  An implementation based on exact prefix sums and binary search has the same distributional guarantees, with an additional logarithmic factor per weighted draw.

The asymptotic bounds in this paper are word-RAM bounds over exact integer values.  If bit complexity is measured, the same algorithms remain exact, and the running time is multiplied by the cost of integer comparisons, additions, indexing, and random generation at the relevant bit lengths.  Since $|J|\le |\mathcal R|\cdot|\mathcal S|$, join weights, atom weights, cumulative weights, and recursive block weights require $O(\log(N+1))$ bits up to constant factors.  For a length-$t$ query, quota counters $H_z$, selected-subtree counters $Q_z$, bucket sizes, and output-coordinate indices require $O(\log(t+1))$ bits.  Thus the discrete arithmetic used by sampling involves $O(\log(N+1)+\log(t+1))$-bit values, in addition to the representation needed for exact endpoint comparisons.

\section{Anchored canonical decomposition}
\label{sec:decomposition}

Fix a suffix instance $J^{(\ell)}(A,B)$ with $\ell<d$.  Dimension $\ell$ is the current ownership dimension.  The side $A$ owns equal-left-endpoint ties: if $L_\ell(a)=L_\ell(b)$, then the pair is assigned to the $A$-anchored case.

\subsection{Anchor later-blocks}

For $a\in A$, define the $B$-side later-block
\[
  X_B^{(\ell)}(a)=
  \left[
    \lb_B^{(\ell)}(L_\ell(a)),
    \lb_B^{(\ell)}(U_\ell(a))
  \right).
\]
This is exactly the contiguous rank interval of all objects $b\in B$ satisfying
\[
  L_\ell(a)\le L_\ell(b)<U_\ell(a).
\]
For $b\in B$, define the $A$-side later-block
\[
  X_A^{(\ell)}(b)=
  \left[
    \ub_A^{(\ell)}(L_\ell(b)),
    \lb_A^{(\ell)}(U_\ell(b))
  \right).
\]
This is exactly the contiguous rank interval of all objects $a\in A$ satisfying
\[
  L_\ell(b)<L_\ell(a)<U_\ell(b).
\]
The lower boundary $\lb_B(L_\ell(a))$ includes equal-left $B$ objects in the $A$-anchored case, while the lower boundary $\ub_A(L_\ell(b))$ excludes equal-left $A$ objects from the $B$-anchored case.  This is the only tie-breaking convention used by the decomposition.

\subsection{Canonical dyadic trees}

For each side $Y\in\{A,B\}$, build a padded full binary segment tree $T_Y^{(\ell)}$ over the positions of $Y^{L_\ell}$.  If $|Y|$ is not a power of two, dummy leaves are appended up to the next power of two.  Dummy leaves never represent objects.  A tree node $z$ represents the real interval
\[
  I_Y(z)=[\lambda(z),\rho(z))\cap[0,|Y|),
\]
where $[\lambda(z),\rho(z))$ is its padded dyadic interval.  Nodes with $I_Y(z)=\emptyset$ are ignored.

For an interval $X=[p,q)\subseteq[0,|Y|)$, define $\Can_Y(X)$ as the set of nodes $z$ satisfying the following conditions:
\begin{enumerate}[leftmargin=1.6em,itemsep=2pt,topsep=2pt]
\item $I_Y(z)\ne\emptyset$;
\item $I_Y(z)\subseteq X$;
\item no ancestor $z'$ of $z$ with $I_Y(z')\ne\emptyset$ satisfies $I_Y(z')\subseteq X$.
\end{enumerate}
Maximality is by tree ancestry.  This convention gives a unique canonical representative even under padding: if two padded nodes induce the same truncated real interval, only the highest one contained in $X$ can be selected.

\begin{lemma}[Canonical cover properties]
\label{lem:canonical-cover}
For every rank interval $X=[p,q)\subseteq[0,|Y|)$, the intervals $\{I_Y(z):z\in\Can_Y(X)\}$ are pairwise disjoint and their union is $X$.  Moreover,
\[
  |\Can_Y(X)|=O(\log(|Y|+1)),
\]
and at every fixed tree depth, the interval $X$ contributes $O(1)$ selected canonical nodes and $O(1)$ unresolved boundary fragments.
\end{lemma}

\begin{proof}
Every position $r\in X$ lies on a unique root-to-leaf path.  The leaf containing $r$ has nonempty real interval contained in $X$, so along that path there is a highest node whose nonempty real interval is contained in $X$.  That node belongs to $\Can_Y(X)$ and covers $r$.  Therefore the selected real intervals cover $X$.

Two selected nodes cannot be in an ancestor-descendant relation, because the descendant would violate maximality.  Incomparable nodes in a binary segment tree have disjoint padded intervals, and hence disjoint real intervals.  Thus the selected real intervals are pairwise disjoint.

At a fixed depth, the dyadic intervals that intersect $X$ form a contiguous run.  Aside from the intervals touching the left and right boundaries of $X$, an interval in this run is either fully contained in $X$ or lies below a node already selected at an earlier depth.  Hence there are only $O(1)$ unresolved boundary fragments at the depth, and only $O(1)$ selected canonical nodes at that depth.  The padded tree height is $O(\log(|Y|+1))$, giving the stated total size.
\end{proof}

\subsection{Tagged local suffix instances}

The two current trees are kept as a tagged disjoint union
\[
  \Znodes_\ell(A,B)=
  (\{\sideB\}\times T_B^{(\ell)})
  \;\sqcup\;
  (\{\sideA\}\times T_A^{(\ell)}).
\]
The tag records which opposite-side tree generated the local instance.  This is part of the mathematical object; a node id from the $A$-tree and a node id from the $B$-tree are never identified.

For a tagged node $z=(\sideB,v)$ in the $B$-tree, define
\[
  A_z^{(\ell)}=
  \{a\in A:v\in\Can_B(X_B^{(\ell)}(a))\},
  \qquad
  B_z^{(\ell)}=
  \{b\in B:\pos_B^{(\ell)}(b)\in I_B(v)\}.
\]
For a tagged node $z=(\sideA,u)$ in the $A$-tree, define
\[
  A_z^{(\ell)}=
  \{a\in A:\pos_A^{(\ell)}(a)\in I_A(u)\},
  \qquad
  B_z^{(\ell)}=
  \{b\in B:u\in\Can_A(X_A^{(\ell)}(b))\}.
\]
The same object may appear in several local nodes, but a join pair appears in exactly one local suffix instance.  The decomposition is pair-disjoint, not object-disjoint.

\begin{lemma}[Prefix certification of local blocks]
\label{lem:prefix-certification}
Assume $(A,B)$ is prefix-certified at level $\ell$.  For every tagged node $z\in\Znodes_\ell(A,B)$, the local instance $(A_z^{(\ell)},B_z^{(\ell)})$ is prefix-certified at level $\ell+1$.
\end{lemma}

\begin{proof}
The certification for dimensions $1,\ldots,\ell-1$ is inherited because $A_z^{(\ell)}\subseteq A$ and $B_z^{(\ell)}\subseteq B$.  It remains to certify dimension $\ell$ for every pair in the local cross product.

For a $B$-tree tagged node $z=(\sideB,v)$, every $a\in A_z^{(\ell)}$ satisfies $v\in\Can_B(X_B^{(\ell)}(a))$, hence $I_B(v)\subseteq X_B^{(\ell)}(a)$.  Every $b\in B_z^{(\ell)}$ satisfies $\pos_B^{(\ell)}(b)\in I_B(v)$.  Therefore
\[
  L_\ell(a)\le L_\ell(b)<U_\ell(a).
\]
Since $b$ is nondegenerate, $L_\ell(b)<U_\ell(b)$, and so $L_\ell(a)<U_\ell(b)$.  Thus $a$ and $b$ intersect in dimension $\ell$.  The $A$-tree case is symmetric: for $z=(\sideA,u)$, every local pair satisfies
\[
  L_\ell(b)<L_\ell(a)<U_\ell(b),
\]
and nondegeneracy of $a$ yields $L_\ell(b)<U_\ell(a)$.  Hence the current dimension is certified for the entire local cross product.
\end{proof}

\begin{theorem}[Anchored canonical decomposition]
\label{thm:decomposition}
For every suffix instance $(A,B)$ and every $\ell<d$,
\[
  J^{(\ell)}(A,B)=
  \biguplus_{z\in\Znodes_\ell(A,B)}
  J^{(\ell+1)}\bigl(A_z^{(\ell)},B_z^{(\ell)}\bigr),
\]
where empty local instances may be omitted.
\end{theorem}

\begin{proof}
Let $(a,b)\in J^{(\ell)}(A,B)$.  Since the pair intersects in dimension $\ell$,
\[
  L_\ell(a)<U_\ell(b),
  \qquad
  L_\ell(b)<U_\ell(a).
\]
If $L_\ell(a)\le L_\ell(b)$, then
\[
  \pos_B^{(\ell)}(b)\in X_B^{(\ell)}(a).
\]
By Lemma~\ref{lem:canonical-cover}, $\Can_B(X_B^{(\ell)}(a))$ partitions $X_B^{(\ell)}(a)$ into disjoint node intervals.  Hence there is a unique node $v\in\Can_B(X_B^{(\ell)}(a))$ whose interval contains $\pos_B^{(\ell)}(b)$.  For the tagged node $z=(\sideB,v)$, we have $a\in A_z^{(\ell)}$ and $b\in B_z^{(\ell)}$, so the pair appears in the $z$-term.  It appears in no other $B$-tree term because the canonical intervals for the fixed anchor $a$ are disjoint.  It appears in no $A$-tree term because $X_A^{(\ell)}(b)$ contains only $A$ objects with left endpoint strictly larger than $L_\ell(b)$.

If $L_\ell(b)<L_\ell(a)$, the symmetric argument assigns the pair to the unique node $u\in\Can_A(X_A^{(\ell)}(b))$ whose interval contains $\pos_A^{(\ell)}(a)$.  The tagged node is $z=(\sideA,u)$.  The pair cannot appear in a $B$-tree term because $X_B^{(\ell)}(a)$ contains only $B$ objects with left endpoint at least $L_\ell(a)$.

Thus every suffix-join pair is assigned exactly once.  Conversely, consider a pair represented by a $B$-tree tagged node $z=(\sideB,v)$: $a\in A_z^{(\ell)}$ and $b\in B_z^{(\ell)}$.  Since $I_B(v)\subseteq X_B^{(\ell)}(a)$ and $\pos_B^{(\ell)}(b)\in I_B(v)$,
\[
  L_\ell(a)\le L_\ell(b)<U_\ell(a).
\]
The box $b$ is nondegenerate in dimension $\ell$, so $L_\ell(b)<U_\ell(b)$, which implies $L_\ell(a)<U_\ell(b)$.  Therefore $a$ and $b$ intersect in dimension $\ell$.  The $A$-tree case gives $L_\ell(b)<L_\ell(a)<U_\ell(b)$ and uses nondegeneracy of $a$ to obtain $L_\ell(b)<U_\ell(a)$.  Each local node therefore certifies the current dimension for every represented cross pair, and the remaining test is exactly the suffix join over dimensions $\ell+1,\ldots,d$.  Lemma~\ref{lem:prefix-certification} records the corresponding recursive invariant for full-dimensional joins.
\end{proof}

\section{One-dimensional terminal atoms}
\label{sec:base}

At level $\ell=d$, the suffix join is one-dimensional and can be represented directly.  The same tie rule is used: the $A$ side owns equal-left-endpoint pairs.

\subsection{Atom family}

For each $a\in A$, define the $A$-anchored atom
\[
  \mathcal A_d(a)=
  \{(a,b):b\in B,\ L_d(a)\le L_d(b)<U_d(a)\}.
\]
Its cardinality is
\[
  \alpha(a)=
  \lb_B^{(d)}(U_d(a))-\lb_B^{(d)}(L_d(a)).
\]
For each $b\in B$, define the $B$-anchored atom
\[
  \mathcal B_d(b)=
  \{(a,b):a\in A,\ L_d(b)<L_d(a)<U_d(b)\}.
\]
Its cardinality is
\[
  \beta(b)=
  \lb_A^{(d)}(U_d(b))-\ub_A^{(d)}(L_d(b)).
\]
Every one-dimensional intersecting pair either satisfies $L_d(a)\le L_d(b)<U_d(a)$ or satisfies $L_d(b)<L_d(a)<U_d(b)$, and the two cases are disjoint.  Therefore
\[
  J^{(d)}(A,B)=
  \left(\biguplus_{a\in A}\mathcal A_d(a)\right)
  \uplus
  \left(\biguplus_{b\in B}\mathcal B_d(b)\right).
\]
Let
\[
  W_d=\sum_{a\in A}\alpha(a)+\sum_{b\in B}\beta(b).
\]
Then $W_d=|J^{(d)}(A,B)|$.

A positive atom is represented by a record
\[
  \tau=(\mathrm{side},\mathrm{anchor},\mathrm{opp},lo,hi,\omega),
\]
where $\mathrm{side}\in\{A,B\}$ indicates the anchor side, $\mathrm{anchor}$ is the anchor object, $\mathrm{opp}$ is a pointer to the terminal-local opposite array sorted by $L_d$, $[lo,hi)$ is the opposite-side rank interval inside that terminal-local array, and $\omega=hi-lo$ is the atom weight.  Write $\Omega_\tau$ for the set of pairs represented by $\tau$.  If $\mathrm{side}=A$, then $\mathrm{opp}=B^{L_d}$ and
\[
  [lo,hi)=[\lb_B(L_d(a)),\lb_B(U_d(a))).
\]
If $\mathrm{side}=B$, then $\mathrm{opp}=A^{L_d}$ and
\[
  [lo,hi)=[\ub_A(L_d(b)),\lb_A(U_d(b))).
\]
The pointer is always to the terminal-local array of the current suffix instance, not to a global $L_d$ view.

\subsection{Computing terminal atoms}

\begin{algorithmbox}{ComputeBaseAtoms$_d(\mathfrak V_d(A,B))$}
\item Initialize temporary arrays $lo_B[a]$, $hi_B[a]$ for $a\in A$ and $lo_A[b]$, $hi_A[b]$ for $b\in B$.
\item Sweep $A^{L_d}$ in increasing $L_d(a)$ while advancing a pointer in $B^{L_d}$ past all objects with $L_d(b)<L_d(a)$; set
\[
  lo_B[a]=\lb_B^{(d)}(L_d(a)).
\]
\item Sweep $A^{U_d}$ in increasing $U_d(a)$ while advancing a pointer in $B^{L_d}$ past all objects with $L_d(b)<U_d(a)$; set
\[
  hi_B[a]=\lb_B^{(d)}(U_d(a)).
\]
\item Sweep $B^{L_d}$ in increasing $L_d(b)$ while advancing a pointer in $A^{L_d}$ past all objects with $L_d(a)\le L_d(b)$; set
\[
  lo_A[b]=\ub_A^{(d)}(L_d(b)).
\]
\item Sweep $B^{U_d}$ in increasing $U_d(b)$ while advancing a pointer in $A^{L_d}$ past all objects with $L_d(a)<U_d(b)$; set
\[
  hi_A[b]=\lb_A^{(d)}(U_d(b)).
\]
\item For every $a\in A$ with $hi_B[a]>lo_B[a]$, emit the atom
\[
  (A,a,B^{L_d},lo_B[a],hi_B[a],hi_B[a]-lo_B[a]).
\]
\item For every $b\in B$ with $hi_A[b]>lo_A[b]$, emit the atom
\[
  (B,b,A^{L_d},lo_A[b],hi_A[b],hi_A[b]-lo_A[b]).
\]
\item Return the atom list $\mathcal T_d$ and its total weight $W_d=\sum_{\tau\in\mathcal T_d}\omega_\tau$.
\end{algorithmbox}

All four rank computations are monotone scans.  ComputeBaseAtoms$_d$ takes $O(|A|+|B|)$ time and $O(|A|+|B|)$ temporary space.  The atom list contains at most $|A|+|B|$ records.

\begin{algorithmbox}{BaseCount$_d(\mathfrak V_d(A,B))$}
\item Run ComputeBaseAtoms$_d(\mathfrak V_d(A,B))$.
\item Return the total atom weight $W_d$.
\end{algorithmbox}

\begin{algorithmbox}{BaseSample$_d(\mathfrak V_d(A,B),h)$}
\item If $h=0$, return the empty sequence.
\item Run ComputeBaseAtoms$_d(\mathfrak V_d(A,B))$ and obtain atoms $\mathcal T_d$ and total weight $W_d$.
\item If $W_d=0$, return \code{EMPTY-INSTANCE}.
\item Build an exact weighted index on the atom weights $(\omega_\tau)_{\tau\in\mathcal T_d}$.
\item For $r=1,\ldots,h$:
  \begin{enumerate}[leftmargin=1.5em,itemsep=1pt,topsep=1pt]
  \item draw an atom $\tau=(\mathrm{side},\mathrm{anchor},\mathrm{opp},lo,hi,\omega)$ with probability $\omega/W_d$;
  \item draw $j$ uniformly from $\{lo,lo+1,\ldots,hi-1\}$;
  \item if $\mathrm{side}=A$, output $(\mathrm{anchor},\mathrm{opp}[j])$; if $\mathrm{side}=B$, output $(\mathrm{opp}[j],\mathrm{anchor})$.
  \end{enumerate}
\item Return the ordered sequence of $h$ pairs.
\end{algorithmbox}

\begin{lemma}[Terminal sampler]
\label{lem:base}
BaseCount$_d$ returns $|J^{(d)}(A,B)|$.  If $h>0$ and $J^{(d)}(A,B)\ne\emptyset$, BaseSample$_d$ returns $h$ independent uniform samples from $J^{(d)}(A,B)$.  If $h>0$ and $J^{(d)}(A,B)=\emptyset$, it returns \code{EMPTY-INSTANCE}.  If $h=0$, it returns the empty sequence.
\end{lemma}

\begin{proof}
The atom families form a disjoint cover of $J^{(d)}(A,B)$, and the weight of each atom is exactly its cardinality.  Therefore BaseCount$_d$ returns the join cardinality.  For a pair $p\in\Omega_\tau$ inside atom $\tau$,
\[
  \Pr[p]=\frac{\omega_\tau}{W_d}\cdot\frac{1}{\omega_\tau}=\frac{1}{W_d}.
\]
The weighted-index draws and the within-atom offset draws are independent across output positions, so the returned sequence is i.i.d. uniform over the one-dimensional join.
\end{proof}

\section{Streaming realization}
\label{sec:streaming}

\ANCHOR-Streaming realizes Theorem~\ref{thm:decomposition} without storing all recursive nodes simultaneously.  At a current level $\ell<d$, the algorithm processes the $B$-tree and the $A$-tree separately.  The two sweeps share the same routing primitive after choosing a source side, a later side, and an orientation map.

\subsection{Orientations and deeper keys}

The $B$-tree orientation has source side $A$ and later side $B$.  A source object $a\in A$ carries the interval $X_B^{(\ell)}(a)$ in the position space of $B^{L_\ell}$, and a node-local callback receives the view $\mathfrak V_{\ell+1}(A_z^{(\ell)},B_z^{(\ell)})$ for a tagged node $z=(\sideB,v)$.

The $A$-tree orientation has source side $B$ and later side $A$.  A source object $b\in B$ carries the interval $X_A^{(\ell)}(b)$ in the position space of $A^{L_\ell}$.  The orientation map swaps the sides before invoking a recursive callback, so every callback still receives a view ordered as $\mathfrak V_{\ell+1}(A_z^{(\ell)},B_z^{(\ell)})$ for a tagged node $z=(\sideA,u)$.

Let
\[
  \mathcal K_{\ell+1}=\{L_{\ell+1},L_{\ell+2},\ldots,L_d,U_d\}
\]
be the set of sorted keys required by the next recursion level.  For every key $K\in\mathcal K_{\ell+1}$, routing maintains source lists and later lists sorted by $K$.

\subsection{Stable routing records}

For one oriented tree, write $S$ for the source side and $Y$ for the later side.  Each later record stores an object $y\in Y$ and its current-position value $\pos_Y^{(\ell)}(y)$.  Each source record stores an object $s\in S$ and its interval $X(s)=[lo(s),hi(s))$ in the position space of $Y^{L_\ell}$.  Empty source intervals are ignored.

For a node $z$, the list $\Lat_K(z)$ contains the later records in $I_Y(z)$, sorted by key $K$.  The list $\Src_K(z)$ contains unresolved source records whose intervals intersect $I_Y(z)$ and have not stopped at an ancestor.  A source record in $\Src_K(z)$ stops at $z$ exactly when
\[
  I_Y(z)\subseteq X(s).
\]
Otherwise it remains unresolved and is sent to every child $c$ with
\[
  I_Y(c)\cap X(s)\ne\emptyset.
\]
A single unresolved source may therefore be copied to both children.  Later records are never copied: each later object belongs to exactly one child interval.  All distributions are stable, meaning that the relative order inherited from a parent $K$-sorted list is preserved inside every produced $K$-sorted list.

This routing rule has a direct canonical interpretation.  For a fixed source object $s$, the nodes where it stops are exactly $\Can_Y(X(s))$; copies that move downward are the unresolved boundary fragments of $X(s)$.

\subsection{RouteTree primitive}

The routing primitive takes both an activity predicate and a request predicate.  The activity predicate determines which subtrees are traversed.  The request predicate determines at which processed nodes a local suffix view is delivered to the callback.

\begin{algorithmbox}{RouteTree$_\ell(S,Y,X,\mathrm{orient},\Active,\Request,\Callback)$}
\item Build the padded segment tree over the positions of $Y^{L_\ell}$; ignore dummy-only nodes.
\item For every key $K\in\mathcal K_{\ell+1}$, initialize the root later list as $Y^K$ and the root source list as the stable restriction of $S^K$ to sources with $X(s)\ne\emptyset$.
\item Process the tree level by level.  A node is present if at least one of its source or later lists is nonempty.  A present node $z$ is processed only if $\Active(z)=\true$.
\item For each processed node $z$:
  \begin{enumerate}[leftmargin=1.5em,itemsep=1pt,topsep=1pt]
  \item For every key $K$, scan $\Src_K(z)$ and form three stable output lists:
  \[
    \Stop_K(z),\qquad \Src_K(\mathrm{left}(z)),\qquad \Src_K(\mathrm{right}(z)).
  \]
  A source is appended to $\Stop_K(z)$ if $I_Y(z)\subseteq X(s)$.  Otherwise it is appended to the left-child list if $I_Y(\mathrm{left}(z))\cap X(s)\ne\emptyset$, and appended to the right-child list if $I_Y(\mathrm{right}(z))\cap X(s)\ne\emptyset$.  Both appends may occur.
  \item For every key $K$, scan $\Lat_K(z)$ and form the stable child lists $\Lat_K(\mathrm{left}(z))$ and $\Lat_K(\mathrm{right}(z))$ by the current-position value.  A later record is appended to exactly one nonempty child interval.
  \item If $\Request(z)=\true$, build the local view from the stopped source lists and the current later lists:
    \begin{itemize}[leftmargin=1.5em,itemsep=1pt,topsep=1pt]
    \item in the $B$-tree orientation, $\Stop_K(z)$ is the $A_z^K$ list and $\Lat_K(z)$ is the $B_z^K$ list;
    \item in the $A$-tree orientation, $\Lat_K(z)$ is the $A_z^K$ list and $\Stop_K(z)$ is the $B_z^K$ list.
    \end{itemize}
    Pass the resulting $\mathfrak V_{\ell+1}(A_z,B_z)$ to $\Callback$.
  \item Retain a child list only if the child has nonempty real interval and $\Active(\mathrm{child})=\true$.
  \end{enumerate}
\item After a level is processed, discard that level's lists and continue with the next level's retained lists.
\end{algorithmbox}

In a counting or node-weight pass, every subtree is active.  The request predicate is true exactly at nodes whose stopped source side and later side are both nonempty; requesting additional zero-size nodes is harmless but unnecessary.  In a sampling pass, a subtree is active exactly when at least one selected output lies in that subtree, and the request predicate is true exactly at nodes with positive local quota.

\begin{lemma}[Local-view invariant]
\label{lem:local-view}
Whenever RouteTree$_\ell$ invokes the callback at a tagged node $z$, the view passed to the callback is exactly
\[
  \mathfrak V_{\ell+1}(A_z^{(\ell)},B_z^{(\ell)}).
\]
For every key $K\in\mathcal K_{\ell+1}$, each local list is sorted by $K$ in the same stable order as the corresponding restricted sorted view.
\end{lemma}

\begin{proof}
The proof is by induction on the depth of the routed tree node.  At the root, the later list for key $K$ is $Y^K$, and the source list is the stable restriction of $S^K$ to sources with nonempty interval.  Both lists are sorted by $K$.

Assume the invariant holds for a node $z$.  The later records in $\Lat_K(z)$ are exactly the objects whose current positions lie in $I_Y(z)$.  Stable distribution by child interval sends each later record to the unique child interval containing its current position.  Thus the child later lists are exactly the restricted later views and remain sorted by $K$.

For a source record $s\in\Src_K(z)$, if $I_Y(z)\subseteq X(s)$, then $s$ has not stopped at an ancestor and $z$ is the highest remaining node on this routed path whose real interval is contained in $X(s)$.  Hence $z\in\Can_Y(X(s))$, and $s$ belongs to $\Stop_K(z)$.  If $I_Y(z)\nsubseteq X(s)$, then $s$ cannot stop at $z$; the unresolved portion of $X(s)$ inside $I_Y(z)$ is represented exactly by the children whose intervals intersect $X(s)$.  Copying $s$ to all such children preserves every boundary fragment.  Since all appends are stable, each stopped or child source list remains sorted by $K$.

Consequently, the stopped source lists at $z$ are precisely the source objects whose canonical decomposition contains the underlying node of $z$, and the current later lists at $z$ are precisely the later objects in that node interval.  The routing predicate for an object depends only on its interval and current-position value, not on the deeper key $K$; hence all stopped lists for different keys contain the same source object set, ordered by their respective keys, and all later lists for different keys contain the same later object set, ordered by their respective keys.  Applying the orientation map yields the tagged local sides $(A_z^{(\ell)},B_z^{(\ell)})$ with all required deeper sorted views.
\end{proof}

\begin{lemma}[Per-depth volume]
\label{lem:per-depth}
For a current suffix instance of size $m=|A|+|B|$, at every fixed depth of either current canonical tree, the total number of routed source records, stopped source records, later records, and node-local records over all nodes at that depth is $O(m)$.  The constant depends only on the fixed dimension $d$.
\end{lemma}

\begin{proof}
Fix one key $K$.  Each later object follows exactly one root-to-leaf path and appears in one later list at the fixed depth.  For a source interval $X=[p,q)$, the routing rule stops a source as soon as a routed node interval is fully contained in $X$.  Therefore the only unresolved routed copies that can survive to a fixed depth are boundary copies whose node intervals intersect the left or right boundary of $X$; there are at most two such copies.  The stopped copies at the same depth are canonical-cover nodes, and Lemma~\ref{lem:canonical-cover} gives at most a constant number of them per depth.  Thus each source object contributes $O(1)$ routed or stopped source records per depth for key $K$.  There are $O(d)$ keys in $\mathcal K_{\ell+1}$, so the total record volume at the depth is $O(m)$.  Node-local records are the stopped source records together with the later records of the same nodes, and therefore satisfy the same bound.
\end{proof}

\subsection{Current-level rank intervals}

Before sweeping the current ownership trees, the algorithm computes every source interval.  For $a\in A$,
\[
  X_B^{(\ell)}(a)=
  [\lb_B^{(\ell)}(L_\ell(a)),\lb_B^{(\ell)}(U_\ell(a))).
\]
For $b\in B$,
\[
  X_A^{(\ell)}(b)=
  [\ub_A^{(\ell)}(L_\ell(b)),\lb_A^{(\ell)}(U_\ell(b))).
\]
The rank values are obtained by binary search in the current opposite left-sorted view.  The cost is $O(m\log(m+1))$ for $m=|A|+|B|$.  When additional upper-endpoint sorted views are maintained, the same ranks can be computed by linear multi-scans, but the binary-search version is sufficient for the stated bounds.

\subsection{Recursive counting}

\begin{algorithmbox}{Count$_\ell(\mathfrak V_\ell(A,B))$}
\item If $A=\emptyset$ or $B=\emptyset$, return $0$.
\item If $\ell=d$, return BaseCount$_d(\mathfrak V_d(A,B))$.
\item Compute all intervals $X_B^{(\ell)}(a)$ for $a\in A$ and $X_A^{(\ell)}(b)$ for $b\in B$.
\item Set $W\leftarrow0$.
\item Run RouteTree$_\ell(A,B,X_B,\Btree,\Active,\Request,\Callback)$ with every subtree active.  For a requested tagged node $z=(\sideB,v)$, the callback computes
\[
  w_z=\mathrm{Count}_{\ell+1}(\mathfrak V_{\ell+1}(A_z,B_z))
\]
and adds $w_z$ to $W$.
\item Run RouteTree$_\ell(B,A,X_A,\Atree,\Active,\Request,\Callback)$ analogously for tagged nodes $z=(\sideA,u)$, and add every recursive weight to $W$.
\item Return $W$.
\end{algorithmbox}

\begin{lemma}[Counting correctness]
\label{lem:counting}
For every suffix instance, Count$_\ell(\mathfrak V_\ell(A,B))$ returns $|J^{(\ell)}(A,B)|$.
\end{lemma}

\begin{proof}
The proof is by induction on the number of remaining dimensions.  The base case is Lemma~\ref{lem:base}.  For $\ell<d$, Lemma~\ref{lem:local-view} supplies every requested node-local view exactly.  By the induction hypothesis, the recursive call at tagged node $z$ returns
\[
  |J^{(\ell+1)}(A_z^{(\ell)},B_z^{(\ell)})|.
\]
Theorem~\ref{thm:decomposition} expresses $J^{(\ell)}(A,B)$ as the tagged disjoint union of these node-local suffix joins.  Summing the recursive weights gives the exact suffix-join cardinality.
\end{proof}

\section{Exact recursive sampling}
\label{sec:sampling}

Sampling uses the same geometric decomposition as counting.  At a non-terminal level, the sampler first computes the exact size of every current node-local suffix join.  It then allocates a multinomial quota over these disjoint pieces, recursively samples only pieces with positive quota, concatenates the block outputs, and shuffles the concatenation.

\subsection{Node weights}

\begin{algorithmbox}{NodeWeights$_\ell(\mathfrak V_\ell(A,B))$}
\item Compute all intervals $X_B^{(\ell)}(a)$ and $X_A^{(\ell)}(b)$.
\item Initialize tagged node-weight arrays $w_z\leftarrow0$ for all $z\in\Znodes_\ell(A,B)$.
\item Run RouteTree$_\ell$ on the $B$-tree with every subtree active and with
\[
  \Request(z)=\true
  \quad\Longleftrightarrow\quad
  \text{the stopped source side and the later side at }z\text{ are both nonempty}.
\]
For every requested tagged node $z=(\sideB,v)$, set
\[
  w_z\leftarrow \mathrm{Count}_{\ell+1}(\mathfrak V_{\ell+1}(A_z,B_z)).
\]
Unrequested nodes retain weight zero.
\item Run RouteTree$_\ell$ on the $A$-tree with the same request predicate, analogously for tagged nodes $z=(\sideA,u)$.
\item Return the interval arrays, tagged node-weight arrays, and $W=\sum_{z\in\Znodes_\ell(A,B)}w_z$.
\end{algorithmbox}

Since each current tree has $O(m)$ nodes, the tagged node-weight arrays use $O(m)$ words.  By Lemma~\ref{lem:counting} and Theorem~\ref{thm:decomposition},
\[
  W=|J^{(\ell)}(A,B)|.
\]

Given a query quota $h$ and node weights $(w_z)_{z\in\Znodes_\ell(A,B)}$ with total $W>0$, the sampler draws
\[
  (H_z)_{z\in\Znodes_\ell(A,B)}
  \sim
  \Mult\left(h;\left(\frac{w_z}{W}\right)_{z\in\Znodes_\ell(A,B)}\right).
\]
For every tree node $z$, define the selected subtree quota
\[
  Q_z=H_z+\sum_{c\text{ child of }z}Q_c,
\]
computed bottom-up inside the same tagged tree.  A node with $H_z=0$ but $Q_z>0$ must still be routed because selected descendants need their local views.  A node with $Q_z=0$ can be skipped together with its entire subtree.

\subsection{Weighted composition lemma}

\begin{lemma}[Weighted quota composition]
\label{lem:quota}
Let $\Omega=\biguplus_{i=1}^k\Omega_i$, let $w_i=|\Omega_i|$, and let $W=\sum_iw_i>0$.  Draw
\[
  (H_1,\ldots,H_k)\sim
  \Mult\left(h;\frac{w_1}{W},\ldots,\frac{w_k}{W}\right).
\]
For each $i$ with $H_i>0$, draw $H_i$ independent uniform samples from $\Omega_i$, independently across all blocks and independently of the quota-generation randomness conditional on the quotas.  Concatenate all generated samples in any deterministic block order and then apply a uniformly random permutation.  The final ordered sequence consists of $h$ independent uniform samples from $\Omega$.
\end{lemma}

\begin{proof}
Consider the coordinate-wise experiment that, independently for every output position, first chooses a label $i$ with probability $w_i/W$ and then draws uniformly from $\Omega_i$.  For any $x\in\Omega_i$, one coordinate equals $x$ with probability
\[
  \frac{w_i}{W}\cdot\frac{1}{w_i}=\frac{1}{W}.
\]
Thus every coordinate is uniform over $\Omega$, and the coordinates are independent.  The vector of label counts in this coordinate-wise experiment has the stated multinomial distribution.  Conditional on those counts, the within-label samples are independent uniform samples inside their respective blocks.  Concatenating by label gives the multiset of labeled coordinate outcomes in a deterministic block order, and a uniform shuffle restores the exchangeable ordered distribution of the coordinate-wise experiment.
\end{proof}

\subsection{Recursive sampler}

\begin{algorithmbox}{Sample$_\ell(\mathfrak V_\ell(A,B),h)$}
\item If $h=0$, return the empty sequence.
\item If $A=\emptyset$ or $B=\emptyset$, return \code{EMPTY-INSTANCE}.
\item If $\ell=d$, return BaseSample$_d(\mathfrak V_d(A,B),h)$.
\item Run NodeWeights$_\ell(\mathfrak V_\ell(A,B))$ and obtain intervals, tagged node weights $(w_z)_z$, and total weight $W$.
\item If $W=0$, return \code{EMPTY-INSTANCE}.
\item Draw multinomial node quotas $(H_z)_z$ from the exact tagged weights $(w_z)_z$.
\item For both current trees, compute selected subtree quotas $Q_z$ bottom-up.
\item Initialize an empty output buffer $\mathcal O$.
\item Run a sample pass over the $B$-tree using RouteTree$_\ell(A,B,X_B,\Btree,\Active,\Request,\Callback)$, where $\Active(z)$ is the predicate $Q_z>0$ and $\Request(z)$ is the predicate $H_z>0$.  On a requested tagged node $z=(\sideB,v)$, append
\[
  \mathrm{Sample}_{\ell+1}(\mathfrak V_{\ell+1}(A_z,B_z),H_z)
\]
to $\mathcal O$.
\item Run the analogous sample pass over the $A$-tree.
\item Apply Fisher--Yates shuffle to $\mathcal O$ and return it.
\end{algorithmbox}

If $H_z>0$, then the exact multinomial draw implies $w_z>0$, so the recursive call is made only on a non-empty local suffix join.  The final shuffle is part of the distributional construction: recursive outputs are produced in implementation order, while the target sequence has no dependence on geometric block order.

\section{Materialized implementations}
\label{sec:materialized}

The same disjoint decomposition supports three materialization levels.  Streaming reconstructs the relevant decomposition on demand.  PrefixCache stores an intermediate frontier.  Compiled stores all terminal atoms and a global index.

\subsection{\texorpdfstring{\ANCHOR-Streaming}{ANCHOR-Streaming}}

The streaming implementation stores the top-level sorted views and uses Count$_1$ or Sample$_1$ directly.  No persistent recursive decomposition is stored.  During one query, recursive views are built as stable lists of object references, used by a callback, and released when that callback returns.  The returned output sequence is not counted as auxiliary space.

\subsection{\texorpdfstring{\ANCHOR-PrefixCache$(k)$}{ANCHOR-PrefixCache(k)}}

For $0\le k\le d-1$, PrefixCache$(k)$ stores the frontier obtained after applying the anchored canonical decomposition through dimensions $1,2,\ldots,k$.  For $k=0$, the frontier consists of the single top-level block.  Let $\mathcal F_k$ be the set of frontier blocks.  For each block $u\in\mathcal F_k$, the data structure stores
\[
  \mathfrak V_{k+1}(A_u,B_u),
  \qquad
  m_u=|A_u|+|B_u|,
  \qquad
  W_u=|J^{(k+1)}(A_u,B_u)|.
\]
Blocks with $W_u=0$ may be removed from the sampling index, although their presence with zero weight does not affect correctness.  Repeated applications of Theorem~\ref{thm:decomposition} give the disjoint union
\[
  J=\biguplus_{u\in\mathcal F_k}J_u,
  \qquad
  J_u=J^{(k+1)}(A_u,B_u).
\]
Preprocessing builds an exact weighted index over the positive frontier weights $W_u$.

A length-$t$ query first handles $t=0$ and the empty-frontier case.  Otherwise, for each output coordinate $r\in\{1,\ldots,t\}$, it independently chooses a frontier block $u_r$ from the weighted index.  Coordinates are then grouped by their selected block.  For every selected block $u$, the streaming suffix sampler is invoked once with quota
\[
  h_u=|\{r:u_r=u\}|.
\]
The returned sequence of length $h_u$ is written into those selected coordinates in increasing coordinate order.  No frontier-level shuffle is used: the frontier choices were made independently for the original output coordinates, so the coordinate order is already the target order.  Suffix sampler calls for different frontier blocks use independent randomness.

The coordinate grouping is implemented sparsely.  A hash table or timestamped dictionary maps a selected frontier-block id to a dynamically growing list of output coordinates, and a separate list records the block ids that have appeared at least once.  No array over all frontier blocks is initialized during a query.  Therefore the grouping and scatter-back work is linear in the number of drawn coordinates and selected block occurrences.

If
\[
  \mathcal F_k^+=\{u\in\mathcal F_k:h_u>0\},
\]
then the query time is
\[
  Q_k(t)=
  O\left(
    t+
    \sum_{u\in\mathcal F_k^+}
      m_u\log^{d-k-1}(m_u+1)
  \right).
\]
The $t$ term accounts for frontier label draws, bucket management, and writing the selected output coordinates.

\subsection{\texorpdfstring{\ANCHOR-Compiled}{ANCHOR-Compiled}}

The compiled implementation expands the anchored decomposition through dimensions $1,2,\ldots,d-1$ and then replaces every terminal one-dimensional instance by its positive ComputeBaseAtoms$_d$ records.  Let $\mathcal T$ be the global list of positive terminal atoms.  Then
\[
  J=\biguplus_{\tau\in\mathcal T}\Omega_\tau,
  \qquad
  \omega_\tau=|\Omega_\tau|.
\]
Each terminal instance stores its local $L_d$-sorted arrays once.  Each atom stores an anchor object, a pointer to the terminal-local opposite array, the rank interval $[lo,hi)$ inside that local array, and the weight $hi-lo$.  Per-atom records do not copy the terminal-local arrays.  A global $L_d$-sorted array is not a valid substitute for the terminal-local opposite array: a globally ranked object may satisfy the last-dimensional rank condition while failing an earlier ownership or intersection condition that defines the terminal instance.

The compiled structure builds one exact weighted index over the atom weights.  If the total atom weight is zero, a positive-length query returns \code{EMPTY-INSTANCE}; a zero-length query returns the empty sequence.  A length-$t$ query independently repeats the following operation for every output coordinate: choose an atom $\tau$ with probability $\omega_\tau/|J|$, draw a uniform offset in $[lo(\tau),hi(\tau))$, and decode the pair using the atom's terminal-local opposite array.  For every $p\in\Omega_\tau$,
\[
  \Pr[p]=\frac{\omega_\tau}{|J|}\cdot\frac{1}{\omega_\tau}=\frac{1}{|J|},
\]
so the coordinates are independent and uniform over $J$.

\begin{center}
\small
\begin{tabular}{@{}>{\raggedright\arraybackslash}p{0.20\linewidth}>{\raggedright\arraybackslash}p{0.35\linewidth}>{\raggedright\arraybackslash}p{0.20\linewidth}>{\raggedright\arraybackslash}p{0.17\linewidth}@{}}
\toprule
Method & Persistent representation & Persistent space & Query time \\
\midrule
Streaming & top-level sorted views only & $O(N)$ auxiliary & $O(N\log^{d-1}(N+1)+t)$ \\
PrefixCache$(k)$ & frontier suffix views and frontier weighted index & $O(N\log^k(N+1))$ & $Q_k(t)$ \\
Compiled & terminal atoms and global weighted index & $O(N\log^{d-1}(N+1))$ & $O(t)$ \\
\bottomrule
\end{tabular}
\end{center}

\section{Correctness and complexity}
\label{sec:guarantees}

\begin{theorem}[Exact i.i.d. correctness]
\label{thm:correctness}
For constant dimension $d$, \ANCHOR-Streaming, \ANCHOR-PrefixCache$(k)$ for every $0\le k\le d-1$, and \ANCHOR-Compiled return an ordered sequence distributed exactly as $t$ independent uniform samples from $J$, under the exact-RAM model.  If $J=\emptyset$ and $t>0$, they return \code{EMPTY-INSTANCE}; if $t=0$, they return the empty sequence.
\end{theorem}

\begin{proof}
For Streaming, use induction on the number of remaining dimensions.  The base case $\ell=d$ is Lemma~\ref{lem:base}.  At a level $\ell<d$, Lemma~\ref{lem:counting} gives exact node weights, and Theorem~\ref{thm:decomposition} partitions $J^{(\ell)}(A,B)$ into tagged pair-disjoint node-local suffix joins.  By the induction hypothesis, every recursive call Sample$_{\ell+1}$ is exact for its local suffix join.  Lemma~\ref{lem:quota} then implies that multinomial quota allocation over exact node weights, recursive sampling inside selected nodes, concatenation, and final shuffling produce an ordered sequence of independent uniform samples from $J^{(\ell)}(A,B)$.

For PrefixCache$(k)$, the stored frontier is obtained by repeated applications of the same tagged disjoint decomposition.  Each output coordinate independently chooses a frontier label with probability proportional to its exact weight and then uses the exact suffix sampler for the chosen frontier block.  The suffix sampler calls use independent randomness across blocks.  Hence every coordinate is uniform over $J$, and the coordinates are independent.

For Compiled, the frontier is refined all the way to terminal one-dimensional atoms.  Atom-proportional sampling followed by a uniform terminal-local offset gives probability $1/|J|$ to every join pair, independently for every output coordinate.  If all exact weights sum to zero, then the corresponding join is empty, and the algorithms return \code{EMPTY-INSTANCE} for positive query length.
\end{proof}

\begin{theorem}[Time and space]
\label{thm:complexity}
Let $N=|\mathcal R|+|\mathcal S|$ and treat $d$ as a constant.  After the top-level sorted views have been constructed, the following bounds hold in the exact-RAM model.
\begin{enumerate}[leftmargin=1.6em,itemsep=2pt,topsep=2pt]
\item \ANCHOR-Streaming answers a length-$t$ query in
\[
  O\bigl(N\log^{d-1}(N+1)+t\bigr)
\]
time and uses $O(N)$ auxiliary space, excluding the returned output sequence.
\item \ANCHOR-PrefixCache$(k)$ preprocesses in $O(N\log^{d-1}(N+1))$ time, uses $O(N\log^k(N+1))$ persistent space, and answers a length-$t$ query in
\[
  O\left(
    t+
    \sum_{u\in\mathcal F_k^+}
      m_u\log^{d-k-1}(m_u+1)
  \right),
\]
where $\mathcal F_k^+$ is the set of frontier blocks selected at least once by the query.
\item \ANCHOR-Compiled preprocesses in $O(N\log^{d-1}(N+1))$ time, uses $O(N\log^{d-1}(N+1))$ persistent space, and answers a length-$t$ query in $O(t)$ time.
\end{enumerate}
For raw unsorted input, add the initial $O(N\log N)$ sorting cost.  This cost is absorbed by the displayed bound when $d\ge2$.  For $d=1$, the algorithmic part after sorted views are available is $O(N+t)$, so the raw-input total is $O(N\log N+t)$.
\end{theorem}

\begin{proof}
Let $C_r(m)$ be the time for Count on a suffix instance of size $m$ with $r$ remaining dimensions.  The base case is $C_1(m)=O(m)$ by the terminal scans.  For $r>1$, computing rank intervals costs $O(m\log(m+1))$.  The two RouteTree sweeps at the current level also cost $O(m\log(m+1))$: by Lemma~\ref{lem:per-depth}, each tree depth has total record volume $O(m)$, and the height is $O(\log(m+1))$.

Let $m_z=|A_z|+|B_z|$ be the local size generated at tagged node $z$ at the current level.  Lemma~\ref{lem:per-depth} gives
\[
  \sum_z m_z=O(m\log(m+1)),
\]
because the total local-record volume is $O(m)$ per depth and there are $O(\log(m+1))$ depths.  Using the induction hypothesis and $m_z\le m$,
\[
\begin{aligned}
  C_r(m)
  &\le O(m\log(m+1))+
      \sum_z O\bigl(m_z\log^{r-2}(m_z+1)\bigr) \\
  &\le O(m\log(m+1))+
      O\bigl(\log^{r-2}(m+1)\sum_zm_z\bigr) \\
  &=O\bigl(m\log^{r-1}(m+1)\bigr).
\end{aligned}
\]
Thus exact counting over the top-level instance takes $O(N\log^{d-1}(N+1))$ time.

Let $S_r(m,h)$ be the time for Sample on a suffix instance of size $m$ with $r$ remaining dimensions and requested length $h$, excluding storage of the returned sequence but including the work needed to write and shuffle it.  For $r=1$, BaseSample takes $O(m+h)$ time.  For $r>1$, Sample first performs NodeWeights, which has the same asymptotic cost as Count.  It then builds the quota distribution over $O(m)$ tagged tree nodes, draws quotas in $O(m+h)$ time, computes subtree quotas in $O(m)$ time, and performs active routing bounded by $O(m\log(m+1))$.  It recurses only on nodes with positive quota.  By the induction hypothesis,
\[
\begin{aligned}
  S_r(m,h)
  &\le O\bigl(m\log^{r-1}(m+1)\bigr)+O(h)
     +\sum_{z:H_z>0}O\bigl(m_z\log^{r-2}(m_z+1)+H_z\bigr) \\
  &\le O\bigl(m\log^{r-1}(m+1)\bigr)+O(h),
\end{aligned}
\]
because $\sum_{z:H_z>0}m_z\le\sum_zm_z=O(m\log(m+1))$ and $\sum_zH_z=h$ at the current level.  At each dimension level, all active recursive calls have quotas summing to $h$, so shuffle and output-buffer work over the full recursion is $O(dh)=O(h)$ for constant $d$.  This proves the Streaming query-time bound.

For Streaming space, RouteTree is implemented level by level, retaining only the current depth's routed buffers and the next depth's buffers.  Lemma~\ref{lem:per-depth} gives $O(m)$ records per depth.  The node-weight, quota, and selected-subtree-quota arrays at a current level also have $O(m)$ entries.  A recursive callback may hold the parent's current-level buffers while processing a child instance, but every child instance has size at most the current parent input size measured as object references in its local view, and the recursion depth is at most $d$.  Since $d$ and the number of sorted views per instance are constant, the auxiliary space is $O(N)$, excluding the output sequence.

For PrefixCache$(k)$, let $M_j$ be the total stored view size after $j$ processed dimensions.  Processing one more dimension replaces each block of size $m_i$ by local blocks of total size $O(m_i\log(m_i+1))\le O(m_i\log(N+1))$.  Therefore
\[
  M_{j+1}\le O(M_j\log(N+1)),
\]
and $M_k=O(N\log^k(N+1))$.  Computing exact frontier weights costs
\[
  \sum_{u\in\mathcal F_k}O\bigl(m_u\log^{d-k-1}(m_u+1)\bigr)
  \le O(N\log^{d-1}(N+1)),
\]
and building the frontier is within the same bound.  During a query, the persistent frontier index draws $t$ labels and stores their output-coordinate positions in sparse buckets.  Since only selected block ids are created and later scanned, this part costs $O(t)$ time and $O(|\mathcal F_k^+|+t)$ transient words.  The suffix streaming sampler is invoked only on selected blocks, giving the displayed adaptive query bound.

For Compiled, the decomposition is expanded through $d-1$ dimensions, so the total terminal-view size is $O(N\log^{d-1}(N+1))$.  Running ComputeBaseAtoms$_d$ on all terminal instances takes time linear in that terminal size, and the global weighted index over positive atoms has the same asymptotic size.  Each output coordinate performs one exact weighted atom draw and one exact uniform offset draw, so a length-$t$ query takes $O(t)$ time in the exact weighted-index model.
\end{proof}

\begin{remark}[Prefix-sum weighted choices]
The distributional statements do not depend on the constant-time weighted-index primitive.  If the compiled atom index or a frontier index is represented by exact prefix sums, each weighted choice can be made by drawing an exact uniform integer in $[0,W)$ and binary searching the prefix array.  The output distribution remains exactly the same.  The corresponding query bounds gain a logarithmic factor in the number of indexed positive weights.
\end{remark}

\section{Implementation invariants}
\label{sec:invariants}

The following invariants give an implementation-level specification of the data represented by every recursive call.
\begin{enumerate}[leftmargin=1.6em,itemsep=2pt,topsep=2pt]
\item Objects are referenced by immutable identifiers.  Sorted views and atom records store references to objects, not copies of coordinate tuples.  Identical boxes remain distinct objects.
\item The recursive suffix instance is always ordered as $(A,B)$.  The $A$-tree orientation uses $B$ as the source side only within the current routing pass; before recursion, the local view is returned to the standard order $(A_z,B_z)$.
\item Source routing uses stable copying.  An unresolved source interval may be copied to both children.  Later routing uses stable splitting, so a later object appears in exactly one child at each depth.
\item Canonical representatives are maximal by tree ancestry among nonempty real intervals contained in the source interval.  Dummy-only nodes are ignored throughout routing, counting, and sampling.
\item A recursive view at level $\ell+1$ contains exactly the deeper keys $L_{\ell+1},\ldots,L_d,U_d$.  The current key $L_\ell$ is not part of the recursive view because dimension $\ell$ has already been certified by the anchor ownership rule.
\item Node-local views passed to callbacks may be temporary buffers or iterators.  In the streaming variant they are released when the callback returns.  PrefixCache and Compiled store only the views or atom tables required by their declared materialization level.
\item In Compiled, each terminal suffix instance stores its terminal-local $A^{L_d}$ and $B^{L_d}$ arrays once.  Atoms store pointers and rank intervals into those arrays.  Decoding an atom against a global $L_d$ array is not valid.
\item Positive-length sampling checks the exact total weight before building or querying a weighted index.  Zero-length sampling returns the empty sequence without consulting the join size.
\item Every recursive sampling call, atom draw, offset draw, quota draw, and shuffle uses fresh randomness, except for the explicit dependence among labels induced by a shared multinomial quota vector.
\end{enumerate}

\section{Summary}
\label{sec:summary}

\ANCHOR{} separates the geometric and probabilistic components of exact join sampling.  Anchored ownership assigns every intersecting cross-color pair to a unique current-dimensional owner.  Canonical dyadic covers turn each anchor's feasible opposite interval into pair-disjoint recursive suffix instances with controlled total size.  Terminal atoms handle the last dimension directly.  Exact integer-weighted quota allocation, recursive sampling, and shuffling produce ordered with-replacement samples from the full join.  Streaming, PrefixCache, and Compiled differ only in how much of the same disjoint decomposition is stored persistently.

\end{document}
