% !TEX program = pdflatex
\documentclass[11pt]{article}

\usepackage[a4paper,margin=1in]{geometry}
\usepackage[T1]{fontenc}
\usepackage[utf8]{inputenc}
\usepackage{lmodern}
\usepackage{microtype}
\usepackage{amsmath,amssymb,amsthm,mathtools}
\usepackage{booktabs}
\usepackage{enumitem}
\usepackage{xcolor}
\usepackage{hyperref}
\usepackage{algorithm}
\usepackage{algpseudocode}

\hypersetup{
  colorlinks=true,
  linkcolor=blue!60!black,
  citecolor=blue!60!black,
  urlcolor=blue!60!black,
  pdftitle={KDS: Exact Uniform Sampling for Half-Open Box-Intersection Joins},
  pdfauthor={}
}

\algrenewcommand\algorithmicrequire{\textbf{Input:}}
\algrenewcommand\algorithmicensure{\textbf{Output:}}
\algrenewcommand\algorithmiccomment[1]{\hfill$\triangleright$ #1}

\theoremstyle{definition}
\newtheorem{definition}{Definition}[section]
\newtheorem{assumption}{Assumption}[section]
\theoremstyle{plain}
\newtheorem{lemma}{Lemma}[section]
\newtheorem{proposition}{Proposition}[section]
\newtheorem{theorem}{Theorem}[section]
\newtheorem{corollary}{Corollary}[section]
\theoremstyle{remark}
\newtheorem{remark}{Remark}[section]

\DeclareMathOperator{\nextUp}{nextUp}
\DeclareMathOperator{\nextDown}{nextDown}
\DeclareMathOperator{\RangeCount}{RangeCount}
\DeclareMathOperator{\CollectItems}{CollectItems}
\DeclareMathOperator{\UniformInt}{UniformInt}
\DeclareMathOperator{\BuildIndex}{BuildIndex}
\DeclareMathOperator{\Draw}{Draw}
\DeclareMathOperator{\payload}{payload}
\DeclareMathOperator{\axis}{axis}
\DeclareMathOperator{\bbox}{bbox}
\DeclareMathOperator{\size}{size}

\newcommand{\Rrel}{\mathcal R}
\newcommand{\Srel}{\mathcal S}
\newcommand{\Jset}{\mathcal J}
\newcommand{\Pset}{\mathcal P}
\newcommand{\Items}{\mathcal I}
\newcommand{\F}{\mathbb F}
\newcommand{\lo}{\mathrm{lo}}
\newcommand{\hi}{\mathrm{hi}}
\newcommand{\pt}{\mathrm{pt}}
\newcommand{\sub}{\mathrm{sub}}
\newcommand{\true}{\mathrm{true}}
\newcommand{\false}{\mathrm{false}}
\newcommand{\nil}{\bot}
\newcommand{\lch}{\mathrm{left}}
\newcommand{\rch}{\mathrm{right}}
\newcommand{\uint}{\mathrm{uint}}
\newcommand{\Real}{\mathbb R}
\newcommand{\Nat}{\mathbb N}

\title{KDS: Exact Uniform Sampling for Half-Open Box-Intersection Joins}
\author{}
\date{}

\begin{document}
\maketitle

\begin{abstract}
KDS is a static sampling algorithm for box-intersection joins.  Given a left relation
$\Rrel$ and a right relation $\Srel$ of $d$-dimensional axis-aligned half-open boxes,
the algorithm returns independent, with-replacement samples that are exactly uniform over
all intersecting pairs in the join result.  The central transformation maps each right-side
box to one point in $2d$ dimensions.  Under strict half-open overlap semantics, a left-side
box induces an orthogonal range query over the lifted point set.  KDS first performs exact
range counting for all left-side boxes, obtaining their join degrees and the total join
cardinality.  A sample is then generated by drawing a left-side box proportionally to its
join degree and drawing a right-side partner uniformly from that box's range-query result.
The right-side conditional draw uses a KD-tree item decomposition that may represent an
entire contained subtree as a single weighted item.  All probability statements in this
document are stated with exact integer random-sampling primitives and explicit floating-point
boundary semantics.
\end{abstract}

\section{Overview}

Let
\[
  \Rrel = \{r_0,r_1,\ldots,r_{n_R-1}\},
  \qquad
  \Srel = \{s_0,s_1,\ldots,s_{n_S-1}\}
\]
be two static relations of axis-aligned boxes in $d$ dimensions.  KDS samples from the box
intersection join
\[
  \Jset = \{(r_i,s_j) : r_i\cap s_j\neq\varnothing\}.
\]
It does not materialize $\Jset$.  Instead, it uses exact join degrees
\[
  w_i = |\{s_j\in\Srel : r_i\cap s_j\neq\varnothing\}|
\]
and a conditional sampler for the matching right-side boxes of a selected left-side box.
The resulting probability factorization is
\[
  \Pr[(i,j)]
  = \Pr[i]\Pr[j\mid i]
  = \frac{w_i}{\sum_h w_h}\cdot \frac{1}{w_i}
  = \frac{1}{|\Jset|}.
\]
The algorithmic work is divided into four phases:
\begin{enumerate}[leftmargin=2em]
  \item lift all right-side boxes into $2d$-dimensional points;
  \item build a static KD-tree over those lifted points;
  \item count the number of matches for every left-side box using closed-interval orthogonal
        range queries;
  \item generate samples by combining exact weighted left-side sampling with exact conditional
        right-side sampling.
\end{enumerate}

The main guarantees are the following.
\begin{enumerate}[leftmargin=2em]
  \item Every output position is distributed exactly uniformly over $\Jset$ whenever
        $|\Jset|>0$.
  \item Output positions are mutually independent and sampling is with replacement.
  \item Boundary-touching boxes are excluded under half-open strict-overlap semantics.
  \item The conversion from strict floating-point inequalities to closed KD-tree ranges is
        exact under the stated finite-coordinate model.
\end{enumerate}

\section{Problem Model}

\subsection{Half-open boxes and strict overlap}

Each box $b$ is represented as
\[
  b = \prod_{a=0}^{d-1} [b^{\lo}_a,b^{\hi}_a),
\]
where $b^{\lo}_a < b^{\hi}_a$ for every dimension $a$.  The intended geometric object is
half-open: the lower boundary is included and the upper boundary is excluded.

\begin{definition}[Strict overlap]
Two boxes $x$ and $y$ strictly overlap if and only if, for every dimension
$a=0,\ldots,d-1$,
\[
  x^{\lo}_a < y^{\hi}_a
  \qquad\text{and}\qquad
  y^{\lo}_a < x^{\hi}_a.
\]
\end{definition}

For half-open intervals, this condition is equivalent to nonempty intersection.  Boundary
contacts of measure zero, such as $x^{\hi}_a=y^{\lo}_a$ in one dimension, do not count as
intersections.

\begin{definition}[Box-intersection join]
The join result is
\[
  \Jset = \{(i,j): 0\le i<n_R,\ 0\le j<n_S,\ r_i\cap s_j\neq\varnothing\}.
\]
Pairs are indexed by array positions.  If two boxes have identical coordinates but different
array positions, their pairs are distinct join-result tuples.
\end{definition}

\subsection{Sampling objective}

Given a target sample size $t\in\Nat$, the required output is a sequence
\[
  Z_0,Z_1,\ldots,Z_{t-1}\in\Jset
\]
such that, when $|\Jset|>0$,
\[
  \Pr[Z_\ell=(i,j)] = \frac{1}{|\Jset|}
\]
for every output position $\ell$ and every $(i,j)\in\Jset$.  The random variables
$Z_0,\ldots,Z_{t-1}$ must be mutually independent.  Therefore the same join pair may appear
multiple times in the output.

If $|\Jset|=0$, there is no valid sample.  The sampling interface returns an empty sequence
and reports that the join cardinality is zero.

\section{Numerical Model}

\subsection{Coordinate domain}

The exact boundary handling of KDS is stated over a fixed floating-point coordinate type
$\F$.  In a concrete implementation, $\F$ is usually IEEE-754 binary32
(\texttt{float32}).  The algorithm is also valid for another fixed finite floating-point
type, provided that all coordinate comparisons, adjacent-number operations, KD-tree bounding
boxes, and query bounds use that same type.

\begin{assumption}[Finite input coordinates]
Every input coordinate is a finite element of $\F$.  NaN and input infinities are excluded.
For every input box $b$ and dimension $a$, $b^{\lo}_a<b^{\hi}_a$ under the numeric comparison
of $\F$.
\end{assumption}

The query representation may use $-\infty$ and $+\infty$ as sentinels, but these values are
not input box coordinates and are not stored as KD-tree point coordinates.

\subsection{Adjacent representable numbers}

For a finite $a\in\F$, let $\nextDown(a)$ be the largest representable value in $\F$ that is
strictly smaller than $a$, and let $\nextUp(a)$ be the smallest representable value in $\F$
that is strictly larger than $a$.  These functions are evaluated in the coordinate type
$\F$.  In a binary32 implementation they correspond to calls such as
\texttt{nextafterf(a, -INFINITY)} and \texttt{nextafterf(a, +INFINITY)}.

\begin{lemma}[Strict-to-closed conversion]
For any finite $a\in\F$ and any $x\in\F$,
\[
  x<a \quad\Longleftrightarrow\quad x\le \nextDown(a),
\]
and
\[
  x>a \quad\Longleftrightarrow\quad x\ge \nextUp(a).
\]
\end{lemma}

\begin{proof}
The representable values of $\F$ form a finite ordered set after excluding NaN.  If
$x<a$, then $x$ is at most the greatest representable value below $a$, namely
$\nextDown(a)$.  Conversely, $x\le\nextDown(a)<a$ implies $x<a$.  The proof for
$\nextUp$ is symmetric.
\end{proof}

Signed zeros, if present in the underlying floating-point type, are compared by their
numeric values.  The adjacent-number operations above should be interpreted using the
implementation's numeric comparison semantics; the equivalences exclude both $+0$ and $-0$
when the strict inequality excludes zero.

\section{Lifting Right-Side Boxes to Points}

Let $D=2d$.  For every right-side box $s_j\in\Srel$, define its lifted point
\[
  p(s_j)
  = \bigl(
      s_{j,0}^{\lo},\ldots,s_{j,d-1}^{\lo},
      s_{j,0}^{\hi},\ldots,s_{j,d-1}^{\hi}
    \bigr) \in \F^{D}.
\]
The first $d$ coordinates store lower endpoints and the last $d$ coordinates store upper
endpoints.  Each lifted point carries payload $j$.

\subsection{Query induced by a left-side box}

For a left-side box $r_i$, the strict-overlap condition with $s_j$ is
\[
  s_{j,a}^{\lo}<r_{i,a}^{\hi}
  \qquad\text{and}\qquad
  s_{j,a}^{\hi}>r_{i,a}^{\lo}
  \qquad\text{for all }a.
\]
Thus the mathematical open query in lifted space is
\[
  \widetilde Q(r_i)
  = \prod_{a=0}^{d-1}(-\infty,r_{i,a}^{\hi})
    \times
    \prod_{a=0}^{d-1}(r_{i,a}^{\lo},+\infty).
\]
The KD-tree implementation uses closed intervals.  The closed query
\[
  Q_i = Q(r_i) = [q_i^{\lo},q_i^{\hi}]
\]
is defined coordinate-wise by
\[
  q_{i,a}^{\lo}=-\infty,
  \qquad
  q_{i,a}^{\hi}=\nextDown(r_{i,a}^{\hi}),
  \qquad a=0,\ldots,d-1,
\]
and
\[
  q_{i,d+a}^{\lo}=\nextUp(r_{i,a}^{\lo}),
  \qquad
  q_{i,d+a}^{\hi}=+\infty,
  \qquad a=0,\ldots,d-1.
\]
A lifted point $p$ matches $Q_i$ if and only if
\[
  q_{i,u}^{\lo}\le p_u\le q_{i,u}^{\hi}
  \qquad\text{for every }u=0,\ldots,D-1.
\]

\begin{lemma}[Lifting equivalence]
For every $r_i\in\Rrel$ and $s_j\in\Srel$,
\[
  p(s_j)\in Q(r_i)
  \quad\Longleftrightarrow\quad
  r_i\cap s_j\neq\varnothing.
\]
\end{lemma}

\begin{proof}
For $a=0,\ldots,d-1$, the closed constraint on coordinate $a$ is
\[
  s_{j,a}^{\lo}\le\nextDown(r_{i,a}^{\hi}),
\]
which is equivalent to $s_{j,a}^{\lo}<r_{i,a}^{\hi}$.  The closed constraint on coordinate
$d+a$ is
\[
  s_{j,a}^{\hi}\ge\nextUp(r_{i,a}^{\lo}),
\]
which is equivalent to $s_{j,a}^{\hi}>r_{i,a}^{\lo}$.  The conjunction of these two
conditions over all dimensions is exactly strict overlap of the two half-open boxes.
\end{proof}

\subsection{Degrees and total join cardinality}

For each left-side box, define its right-side match set and degree as
\[
  S(r_i) = \{s_j\in\Srel : r_i\cap s_j\neq\varnothing\},
  \qquad
  w_i = |S(r_i)|.
\]
By the lifting equivalence,
\[
  w_i = |\{p(s_j):p(s_j)\in Q(r_i)\}|.
\]
The join result decomposes as a disjoint union over left-side indices:
\[
  \Jset = \biguplus_{i=0}^{n_R-1}\{i\}\times S(r_i),
\]
and therefore
\[
  |\Jset| = \sum_{i=0}^{n_R-1} w_i.
\]
KDS computes every $w_i$ exactly and stores
\[
  W = \sum_i w_i.
\]

\section{Exact Random-Sampling Primitives}

The correctness theorems use exact integer sampling primitives.  A deterministic pseudo-random
number generator may be used in software, but the mathematical guarantees are stated for the
ideal distribution generated by the following operations.

\subsection{Uniform bounded integers}

\begin{definition}[Exact bounded integer draw]
For an integer $M>0$, $\UniformInt(0,M-1)$ returns each value in
$\{0,1,\ldots,M-1\}$ with probability exactly $1/M$.
\end{definition}

A standard implementation uses rejection sampling from fixed-width random words.  If one
word has range $0,\ldots,2^b-1$, set
\[
  L = \left\lfloor\frac{2^b}{M}\right\rfloor M.
\]
Draw a word $X$.  If $X<L$, return $X\bmod M$; otherwise, reject and draw again.  If
$M>2^b$, concatenate independent words until the temporary integer range covers $M$.  This
avoids modulo bias.

\subsection{Weighted sampling by prefix sums}

\begin{definition}[Exact weighted draw]
Given nonnegative integer weights
$\omega_0,\ldots,\omega_{k-1}$ with
$\Omega=\sum_{h=0}^{k-1}\omega_h>0$, an exact weighted draw returns label $h$ with
probability $\omega_h/\Omega$.  Zero-weight labels are never returned.
\end{definition}

The prefix-sum construction stores
\[
  C_h=\sum_{g=0}^{h}\omega_g.
\]
A draw samples $U\sim\UniformInt(0,\Omega-1)$ and returns the unique $h$ satisfying
\[
  C_{h-1}\le U<C_h,
  \qquad C_{-1}=0.
\]
This method is exact.  Its build cost is $O(k)$, storage is $O(k)$, and draw cost is
$O(\log k)$ with binary search or $O(1)$ only in special cases where direct indexing is
available.

\subsection{Weighted sampling by an integer alias table}

For applications requiring $O(1)$ draw time after preprocessing, KDS may use an integer alias
index.  The construction stores, for each column $c\in\{0,\ldots,k-1\}$, a primary label
$a_c$, an alias label $b_c$, and an integer threshold $\tau_c\in\{0,\ldots,\Omega\}$ such
that for every label $h$,
\[
  \sum_{c:a_c=h}\tau_c + \sum_{c:b_c=h}(\Omega-\tau_c) = k\omega_h.
\]
A draw samples
\[
  C\sim\UniformInt(0,k-1),
  \qquad
  U\sim\UniformInt(0,\Omega-1),
\]
and returns $a_C$ if $U<\tau_C$, otherwise $b_C$.

\begin{lemma}[Exactness of the integer alias draw]
If the alias-table identity above holds, the draw returns each label $h$ with probability
$\omega_h/\Omega$.
\end{lemma}

\begin{proof}
The contribution of column $c$ to label $h$ is $\tau_c/\Omega$ if $a_c=h$ and
$(\Omega-\tau_c)/\Omega$ if $b_c=h$.  Since each column is selected with probability $1/k$,
\[
  \Pr[h]
  = \frac{1}{k\Omega}
      \left(
        \sum_{c:a_c=h}\tau_c
        +
        \sum_{c:b_c=h}(\Omega-\tau_c)
      \right)
  = \frac{k\omega_h}{k\Omega}
  = \frac{\omega_h}{\Omega}.
\]
\end{proof}

A standard alias construction realizes this identity by treating column capacity as
$\Omega$ and label mass as $k\omega_h$.  Labels with mass below $\Omega$ are placed on a
small stack and labels with mass at least $\Omega$ are placed on a large stack.  Pairing a
small label with a large label creates one column.  All arithmetic is integer arithmetic.
The products $k\omega_h$ must be computed without overflow.  If the available fixed-width
integer type cannot safely represent these products, the implementation uses a wider integer
type, arbitrary-precision integers, or falls back to the prefix-sum index.  The correctness
proof is independent of the selected representation as long as all integer masses are exact.

\subsection{Weighted-index interface}

The algorithms below use an abstract weighted-index interface:
\[
  A = \BuildIndex(\omega_0,\ldots,\omega_{k-1}),
  \qquad
  h = \Draw(A).
\]
The index may be a prefix-sum index or an integer alias index.  The probability guarantees
require only exact weighted sampling.  The time bounds explicitly state which draw cost is
being used.

\section{KD-Tree over Lifted Points}

\subsection{Stored arrays and payloads}

Let $P[0:n_S)$ be the array of lifted points.  Each entry stores the $D$ coordinates of
$p(s_j)$ and the payload $j$.  KD-tree construction may rearrange $P$ in place.  The payload
field makes the right-side object identifier invariant under rearrangement.

\subsection{Node fields}

Each nonempty KD-tree node $v$ represents a contiguous interval
\[
  [l_v,r_v)\subseteq [0,n_S)
\]
of the current point array.  Its median position is
\[
  m_v = l_v + \left\lfloor\frac{r_v-l_v}{2}\right\rfloor.
\]
The node stores:
\begin{enumerate}[leftmargin=2em]
  \item its left child $\lch(v)$ and right child $\rch(v)$;
  \item its interval endpoints $l_v$ and $r_v$;
  \item its median position $m_v$;
  \item its split dimension $\axis(v)=\operatorname{depth}(v)\bmod D$;
  \item its subtree bounding box
        $\bbox(v)=[\lo(v),\hi(v)]\in\F^D\times\F^D$.
\end{enumerate}
The subtree size is
\[
  \size(v)=r_v-l_v.
\]
For a node $v$, let $P(v)$ denote the set of point-array entries in its subtree.

\subsection{Construction algorithm}

\begin{algorithm}[htbp]
\caption{\textsc{BuildKDTree}$(P,l,r,\delta)$}
\begin{algorithmic}[1]
\Require lifted point array $P$, interval $[l,r)$, recursion depth $\delta$
\Ensure root node of the KD-tree over $P[l:r)$, or $\nil$ when $l\ge r$
\If{$l\ge r$}
  \State \Return $\nil$
\EndIf
\State $a\gets \delta\bmod D$
\State $m\gets l+\lfloor(r-l)/2\rfloor$
\State partition $P[l:r)$ by the $a$-th coordinate so that $P[m]$ is a median element
\State create node $v$ with interval $[l,r)$, median position $m$, and split axis $a$
\State $\lch(v)\gets\textsc{BuildKDTree}(P,l,m,\delta+1)$
\State $\rch(v)\gets\textsc{BuildKDTree}(P,m+1,r,\delta+1)$
\State compute $\bbox(v)$ from $P[m]$, $\bbox(\lch(v))$, and $\bbox(\rch(v))$
\State \Return $v$
\end{algorithmic}
\end{algorithm}

Ties in the selected coordinate may be broken by any deterministic total order over point
records, such as lexicographic order over coordinates followed by payload.  The correctness
of range counting and sampling does not depend on the tie-breaking rule.

\begin{lemma}[KD-tree invariants]
After construction, the following statements hold for every nonempty node $v$.
\begin{enumerate}[leftmargin=2em]
  \item The subtree of $v$ is represented by the contiguous interval $[l_v,r_v)$.
  \item The point sets associated with $P[m_v]$, $\lch(v)$, and $\rch(v)$ are pairwise
        disjoint and their union is $P(v)$.
  \item Every point appears as the median point of exactly one KD-tree node.
  \item The bounding box $\bbox(v)$ is exact: for every coordinate $u$,
        \[
          \lo_u(v)=\min_{p\in P(v)}p_u,
          \qquad
          \hi_u(v)=\max_{p\in P(v)}p_u.
        \]
\end{enumerate}
\end{lemma}

\begin{proof}
The construction splits the interval $[l,r)$ into $[l,m)$, the singleton $\{m\}$, and
$[m+1,r)$.  The recursive calls operate on the two disjoint subintervals.  Hence the
contiguous-interval and disjoint-union properties follow by induction.  Since every recursive
step removes exactly one median point from the interval and recurses on the remaining two
parts, each point is chosen as a median exactly once.  The bounding box is computed by
coordinate-wise minimum and maximum over the median point and the two child bounding boxes,
which is exact by the induction hypothesis.
\end{proof}

\section{Exact Range Counting}

\subsection{Bounding-box predicates}

For a closed query $Q=[q^{\lo},q^{\hi}]$, a subtree bounding box $\bbox(v)$ is disjoint
from $Q$ when
\[
  \exists u\in\{0,\ldots,D-1\}:
  \hi_u(v)<q^{\lo}_u\ \text{ or }\ \lo_u(v)>q^{\hi}_u.
\]
It is contained in $Q$ when
\[
  \forall u\in\{0,\ldots,D-1\}:
  q^{\lo}_u\le\lo_u(v)\ \text{ and }\ \hi_u(v)\le q^{\hi}_u.
\]
A point $p$ matches $Q$ when
\[
  \forall u\in\{0,\ldots,D-1\}: q^{\lo}_u\le p_u\le q^{\hi}_u.
\]

\subsection{Counting algorithm}

\begin{algorithm}[htbp]
\caption{\textsc{RangeCount}$(v,Q)$}
\begin{algorithmic}[1]
\Require KD-tree node $v$ and closed query $Q=[q^{\lo},q^{\hi}]$
\Ensure number of points in subtree $v$ that match $Q$
\If{$v=\nil$}
  \State \Return $0$
\EndIf
\If{$\bbox(v)$ is disjoint from $Q$}
  \State \Return $0$
\EndIf
\If{$\bbox(v)\subseteq Q$}
  \State \Return $r_v-l_v$
\EndIf
\State $c\gets 0$
\If{$P[m_v]\in Q$}
  \State $c\gets c+1$
\EndIf
\State $c\gets c+\textsc{RangeCount}(\lch(v),Q)$
\State $c\gets c+\textsc{RangeCount}(\rch(v),Q)$
\State \Return $c$
\end{algorithmic}
\end{algorithm}

\begin{lemma}[Range-count correctness]
For every KD-tree node $v$ and every closed query $Q$, $\RangeCount(v,Q)$ returns
$|P(v)\cap Q|$.
\end{lemma}

\begin{proof}
The proof is by induction on subtree size.  The empty subtree is immediate.  If
$\bbox(v)$ is disjoint from $Q$, exactness of the bounding box implies that no point in
$P(v)$ can match $Q$.  If $\bbox(v)\subseteq Q$, all points in the subtree match $Q$, and the
correct count is $\size(v)=r_v-l_v$.  In the remaining case, the KD-tree invariants give the
pairwise-disjoint decomposition
\[
  P(v)=\{P[m_v]\}\uplus P(\lch(v))\uplus P(\rch(v)).
\]
The algorithm counts the median point directly and applies the induction hypothesis to the
two child subtrees.  Summing these three disjoint contributions gives $|P(v)\cap Q|$.
\end{proof}

\begin{corollary}[Degree correctness]
For each left-side box $r_i$, if $Q_i=Q(r_i)$, then
\[
  \RangeCount(T,Q_i)=w_i.
\]
\end{corollary}

\begin{proof}
By range-count correctness, $\RangeCount(T,Q_i)$ is the number of lifted points in $Q_i$.
By the lifting equivalence, those points are exactly the right-side boxes intersecting
$r_i$.
\end{proof}

\section{Conditional Sampling by Item Decomposition}

\subsection{Item types}

After the left-side index $i$ has been selected, KDS needs a strictly uniform draw from
$S(r_i)$.  The range-query result can be represented by an item collection rather than by an
explicit list of every matching right-side point.

There are two item types.
\begin{enumerate}[leftmargin=2em]
  \item A \emph{point item} stores one matching point-array position $h$ and has weight $1$.
  \item A \emph{subtree item} stores a node $v$ whose bounding box is fully contained in the
        query and has weight $\size(v)=r_v-l_v$.
\end{enumerate}
In adversarial configurations, the decomposition may contain one point item per matching
point.  In favorable configurations, a large contained subtree is represented by one item.
In all cases, the complete join result $\Jset$ is never materialized.

\subsection{Collection algorithm}

\begin{algorithm}[htbp]
\caption{\textsc{CollectItems}$(v,Q,\Items)$}
\begin{algorithmic}[1]
\Require KD-tree node $v$, closed query $Q$, appendable list $\Items$
\Ensure append a disjoint item representation of $P(v)\cap Q$
\If{$v=\nil$}
  \State \Return
\EndIf
\If{$\bbox(v)$ is disjoint from $Q$}
  \State \Return
\EndIf
\If{$v$ is a leaf}
  \State append point item $(\pt,m_v)$ to $\Items$
  \State \Return
\EndIf
\If{$P[m_v]\in Q$}
  \State append point item $(\pt,m_v)$ to $\Items$
\EndIf
\For{each child $c\in\{\lch(v),\rch(v)\}$}
  \If{$c=\nil$}
    \State continue
  \ElsIf{$\bbox(c)$ is disjoint from $Q$}
    \State continue
  \ElsIf{$\bbox(c)\subseteq Q$}
    \State append subtree item $(\sub,c)$ to $\Items$
  \Else
    \State \textsc{CollectItems}$(c,Q,\Items)$
  \EndIf
\EndFor
\end{algorithmic}
\end{algorithm}

The leaf case appends the leaf point without a separate point-in-query test.  This is safe
because a leaf bounding box is the point itself; after the disjointness test has failed, that
point must lie in the closed query.

\subsection{Coverage and disjointness}

For an item $I$, define its contribution set as
\[
  \Gamma(I)=
  \begin{cases}
    \{P[h]\}, & I=(\pt,h),\\
    P(v), & I=(\sub,v).
  \end{cases}
\]
Its weight is
\[
  \omega(I)=|\Gamma(I)|=
  \begin{cases}
    1, & I=(\pt,h),\\
    r_v-l_v, & I=(\sub,v).
  \end{cases}
\]

\begin{lemma}[Item coverage and disjointness]
Let $\Items(Q)$ be the item collection produced by $\CollectItems(T,Q,\Items)$.  Then
\[
  P(T)\cap Q = \biguplus_{I\in\Items(Q)} \Gamma(I).
\]
Consequently,
\[
  \sum_{I\in\Items(Q)}\omega(I)=|P(T)\cap Q|.
\]
\end{lemma}

\begin{proof}
At each internal node, the KD-tree invariants decompose the subtree into the median point,
the left child, and the right child, with no overlap.  The median point is appended if and
only if it matches the query.  For each child subtree, the algorithm either discards it when
its bounding box is disjoint from the query, appends the whole child as a subtree item when
its bounding box is contained in the query, or recursively applies the same rule.  A subtree
that is appended as an item is not visited again, so no point in it can be appended elsewhere.
Every non-discarded, non-contained subtree is eventually decomposed until its matching points
are represented.  Thus the item contribution sets form a disjoint union of exactly the
matching points.
\end{proof}

\subsection{Conditional draw from items}

Given $\Items(Q_i)$, build an exact weighted index over item weights $\omega(I)$.  To draw
one right-side partner:
\begin{enumerate}[leftmargin=2em]
  \item draw an item $I$ with probability $\omega(I)/\sum_{I'}\omega(I')$;
  \item if $I=(\pt,h)$, return $\payload(P[h])$;
  \item if $I=(\sub,v)$, draw $H\sim\UniformInt(l_v,r_v-1)$ and return $\payload(P[H])$.
\end{enumerate}
Because a subtree item is emitted only when its entire bounding box is contained in the query,
every point-array position $H\in[l_v,r_v)$ corresponds to a valid matching right-side box.

\begin{lemma}[Conditional uniformity]
Fix $i$ with $w_i>0$.  The conditional item sampler for $Q_i=Q(r_i)$ returns every
$s_j\in S(r_i)$ with probability exactly $1/w_i$.
\end{lemma}

\begin{proof}
By item coverage and disjointness, every matching lifted point lies in exactly one item, and
the total item weight is $w_i$.  If the point lies in a point item, the probability of
selecting that item is $1/w_i$.  If the point lies in a subtree item of size $m$, the item is
selected with probability $m/w_i$, and the point is then selected from the subtree interval
with probability $1/m$.  The product is $1/w_i$.
\end{proof}

\section{Complete KDS State and Algorithms}

\subsection{State contents}

A KDS state contains:
\begin{enumerate}[leftmargin=2em]
  \item the input dimension $d$ and lifted dimension $D=2d$;
  \item a reference to the left-side box array $\Rrel$, or a cached array of queries
        $Q_0,\ldots,Q_{n_R-1}$;
  \item the lifted point array $P$ with right-side payloads;
  \item the KD-tree root $T$ and node array;
  \item the degree array $w_0,\ldots,w_{n_R-1}$;
  \item the total join cardinality $W=\sum_i w_i$;
  \item an exact weighted index $A_R$ over the left-side degrees when $W>0$.
\end{enumerate}
If queries are not cached, each $Q_i$ is reconstructed from $r_i$ when needed.  Both choices
produce the same distribution.  Caching uses $O(Dn_R)$ additional persistent memory and saves
query-bound reconstruction time.

\subsection{Build and count}

\begin{algorithm}[htbp]
\caption{\textsc{BuildKDS}$(\Rrel,\Srel,d)$}
\begin{algorithmic}[1]
\Require left-side boxes $\Rrel$, right-side boxes $\Srel$, dimension $d$
\Ensure KDS state
\State validate that all input coordinates are finite and satisfy $\lo<\hi$ in each dimension
\State $D\gets 2d$
\For{$j=0$ to $n_S-1$}
  \State $P[j]\gets p(s_j)$ and set $\payload(P[j])\gets j$
\EndFor
\State $T\gets\textsc{BuildKDTree}(P,0,n_S,0)$
\For{$i=0$ to $n_R-1$}
  \State $Q_i\gets Q(r_i)$
  \State $w_i\gets\textsc{RangeCount}(T,Q_i)$
\EndFor
\State $W\gets \sum_{i=0}^{n_R-1}w_i$ using an overflow-checked accumulator
\If{$W>0$}
  \State $A_R\gets\BuildIndex(w_0,\ldots,w_{n_R-1})$
\Else
  \State $A_R\gets\nil$
\EndIf
\State \Return state $(d,D,\Rrel,P,T,w,W,A_R)$
\end{algorithmic}
\end{algorithm}

\subsection{Single sample}

\begin{algorithm}[htbp]
\caption{\textsc{SampleOne}$(\mathrm{state})$}
\begin{algorithmic}[1]
\Require KDS state
\Ensure one join pair $(i,j)$, or EMPTY if $W=0$
\If{$W=0$}
  \State \Return EMPTY
\EndIf
\State $i\gets\Draw(A_R)$
\State $Q_i\gets Q(r_i)$
\State $\Items\gets$ empty list
\State \textsc{CollectItems}$(T,Q_i,\Items)$
\State $A_i\gets\BuildIndex(\{\omega(I):I\in\Items\})$
\State $I\gets\Draw(A_i)$
\If{$I=(\pt,h)$}
  \State \Return $(i,\payload(P[h]))$
\Else
  \State let $I=(\sub,v)$
  \State $H\gets\UniformInt(l_v,r_v-1)$
  \State \Return $(i,\payload(P[H]))$
\EndIf
\end{algorithmic}
\end{algorithm}

\begin{theorem}[Single-sample uniformity]
Assume $W=|\Jset|>0$.  \textsc{SampleOne} returns every pair $(i,j)\in\Jset$ with
probability exactly $1/W$.
\end{theorem}

\begin{proof}
For any $(i,j)\in\Jset$, $s_j\in S(r_i)$ and $w_i>0$.  The left-side draw returns $i$ with
probability $w_i/W$.  Conditional on $i$, the item sampler returns $j$ with probability
$1/w_i$.  Therefore
\[
  \Pr[(i,j)] = \frac{w_i}{W}\cdot\frac{1}{w_i} = \frac{1}{W}.
\]
\end{proof}

\subsection{Batch sampling with left-side grouping}

Batch sampling avoids rebuilding the same conditional item index many times when the same
left-side index appears repeatedly.  It first samples all left-side indices independently,
then groups output slots by the sampled left-side index.

\begin{algorithm}[htbp]
\caption{\textsc{BatchSample}$(\mathrm{state},t)$}
\begin{algorithmic}[1]
\Require KDS state and sample size $t$
\Ensure sequence $Z[0:t)$ of join-pair samples, or the empty sequence if $t=0$ or $W=0$
\If{$t=0$ or $W=0$}
  \State \Return empty sequence
\EndIf
\State $L\gets$ empty list
\For{$\ell=0$ to $t-1$}
  \State $i_\ell\gets\Draw(A_R)$
  \State append $(i_\ell,\ell)$ to $L$
\EndFor
\State group $L$ by left-side index, using sorting, hashing, counting buckets, or radix buckets
\State initialize output array $Z[0:t)$
\For{each group $G_i=\{(i,\ell):\ell\text{ has sampled left index }i\}$}
  \State $Q_i\gets Q(r_i)$
  \State $\Items\gets$ empty list
  \State \textsc{CollectItems}$(T,Q_i,\Items)$
  \State $A_i\gets\BuildIndex(\{\omega(I):I\in\Items\})$
  \For{each $(i,\ell)\in G_i$}
    \State $I\gets\Draw(A_i)$
    \If{$I=(\pt,h)$}
      \State $j\gets\payload(P[h])$
    \Else
      \State let $I=(\sub,v)$
      \State $H\gets\UniformInt(l_v,r_v-1)$
      \State $j\gets\payload(P[H])$
    \EndIf
    \State $Z[\ell]\gets(i,j)$
  \EndFor
\EndFor
\State \Return $Z$
\end{algorithmic}
\end{algorithm}

\begin{theorem}[Batch-sampling correctness]
If $W>0$, then \textsc{BatchSample}$(\mathrm{state},t)$ returns $t$ mutually independent,
with-replacement samples, each exactly uniform over $\Jset$.
\end{theorem}

\begin{proof}
The first phase performs $t$ independent draws from the exact left-side index $A_R$, so
$\Pr[i_\ell=i]=w_i/W$ independently for every output slot $\ell$.  Conditioning on the
sampled left-side indices, the grouped phase processes slots in an arbitrary order but writes
results back to their original positions.  For every slot with left-side index $i$, the
right-side draw uses fresh independent random choices from the exact conditional item index
for $Q_i$.  By conditional uniformity, the returned right-side index is uniform over
$S(r_i)$.  Therefore each slot has the same distribution as \textsc{SampleOne}, and the
fresh random choices across slots preserve mutual independence.  Grouping changes only the
order of computation, not the random variables assigned to output positions.
\end{proof}

\begin{theorem}[End-to-end correctness]
Under the finite-coordinate model and exact integer random primitives, KDS returns strictly
uniform, independent, with-replacement samples over the complete half-open box-intersection
join result.
\end{theorem}

\begin{proof}
The lifting equivalence maps right-side partners of every left-side box exactly to a
closed-interval KD-tree range query.  Range counting computes every degree $w_i$ exactly, so
$W=\sum_iw_i=|\Jset|$.  The left-side weighted draw selects $i$ with probability $w_i/W$.
The conditional item sampler selects every $s_j\in S(r_i)$ with probability $1/w_i$.  The
single-sample and batch-sampling theorems then give the desired distribution and independence.
\end{proof}

\section{Complexity}

Let
\[
  n=n_R,
  \qquad
  m=n_S,
  \qquad
  D=2d.
\]
The dimension factor $D$ is shown explicitly.  In the common implementation range
$d\in\{2,3,4,5\}$, $D\in\{4,6,8,10\}$ is a small constant, and the same bounds may be read
with $D$ suppressed.

\subsection{KD-tree construction}

A median-split KD-tree has $m$ nonempty nodes.  If median selection uses linear-time
partitioning at each recursion level, the typical number of coordinate comparisons is
$O(m\log m)$.  Building exact node bounding boxes costs $O(Dm)$.  If complete $D$-coordinate
point records are moved during partitioning, the data-movement cost may be written as
$O(Dm\log m)$; if the implementation partitions an index array pointing to immutable
coordinate storage, the movement cost is $O(m\log m)$ and the coordinate storage is
$O(Dm)$.

The persistent KD-tree storage is
\[
  O(Dm+m),
\]
including lifted coordinates, payloads, child links, interval endpoints, split axes, and
$2D$ bounding-box coordinates per node.  When $D$ is fixed, this is $O(m)$.

\subsection{Counting phase}

Let $V_{\mathrm{count}}(r_i)$ be the number of KD-tree nodes visited by
$\RangeCount(T,Q_i)$.  Each visited node performs a constant number of $D$-dimensional
bounding-box or point comparisons.  Hence the counting phase costs
\[
  O\left(D\sum_{i=0}^{n-1}V_{\mathrm{count}}(r_i)\right)
\]
for the tree queries, plus
\[
  O(n)
\]
for summing the degree array and constructing a left-side weighted index with linear build
cost.

KD-trees do not give a stable polylogarithmic worst-case bound for arbitrary orthogonal range
queries in the lifted dimension.  In the worst case, a query may visit $\Theta(m)$ nodes, and
the full counting phase may take $\Theta(Dnm)$ time.  The visited-node expression is the
operational cost relevant to a given input distribution and query workload.

\subsection{Conditional item construction}

For a selected left-side box $r_i$, let
\[
  V_{\mathrm{item}}(r_i)
\]
be the number of nodes visited by $\CollectItems(T,Q_i)$ and let
\[
  B(r_i)=|\Items(Q_i)|
\]
be the number of generated items.  Item construction costs
\[
  O(DV_{\mathrm{item}}(r_i)+B(r_i)).
\]
Building a conditional weighted index over the items costs $O(B(r_i))$ for either prefix sums
or a linear-time alias construction.  In the worst case,
\[
  V_{\mathrm{item}}(r_i)=\Theta(m),
  \qquad
  B(r_i)=\Theta(w_i),
\]
and $B(r_i)$ may be as large as $m$.

\subsection{Draw costs under different weighted indexes}

Let $T_L$ be the cost of drawing one left-side index from $A_R$, and let
$T_R(B)$ be the cost of drawing one item from a conditional item index of size $B$.
\begin{center}
\begin{tabular}{lll}
\toprule
Index type & Build cost & Draw cost \\
\midrule
Prefix sums + binary search & $O(k)$ & $O(\log k)$ \\
Integer alias table & $O(k)$ & $O(1)$ \\
\bottomrule
\end{tabular}
\end{center}
Thus, with integer alias indexes for both $A_R$ and conditional item indexes, the additional
cost per sampled output after the corresponding indexes are built is $O(1)$.  With prefix
indexes, the logarithmic search terms remain in the bound.

\subsection{Batch-sampling time}

Let $R^\star$ be the set of distinct left-side indices sampled in the batch, and let $k_i$
be the number of output slots in the group for $i$.  Then
\[
  \sum_{i\in R^\star}k_i=t.
\]
Let $G(t,n)$ be the grouping cost.  Sorting gives $G(t,n)=O(t\log t)$.  Hash grouping gives
expected $O(t)$ under the hash model.  Counting buckets give $O(t+n)$ time and $O(n)$ bucket
space.  Radix grouping gives near-linear time for fixed-width integer keys.

A general batch-sampling bound is
\[
  O\left(
      tT_L + G(t,n)
      + \sum_{i\in R^\star}
          \bigl(DV_{\mathrm{item}}(r_i)+B(r_i)+k_iT_R(B(r_i))\bigr)
    \right).
\]
With integer alias indexes and hash or radix grouping, this simplifies to
\[
  O\left(
      t
      + G(t,n)
      + \sum_{i\in R^\star}\bigl(DV_{\mathrm{item}}(r_i)+B(r_i)+k_i\bigr)
    \right).
\]

\subsection{Space}

Persistent state uses
\[
  O(Dm+m+n)
\]
space for lifted points, the KD-tree, the degree array, and the left-side weighted index.
If all left-side queries are cached, add $O(Dn)$ space.  Temporary sampling space consists
of the grouping records, the output array, and the current item list and weighted index.  A
single batch of size $t$ uses $O(t)$ output and grouping space.  Chunked sampling with chunk
size $c$ limits this materialized space to $O(c)$, aside from any externally stored final
output.

\section{Implementation Details}

\subsection{Input validation}

A concrete implementation should check the following conditions before building the index.
\begin{enumerate}[leftmargin=2em]
  \item The dimension $d$ is positive and is within the implementation-supported range.
  \item Every coordinate is finite in the selected coordinate type.
  \item No coordinate is NaN.
  \item For every box and every dimension, $\lo<\hi$ under the coordinate type's comparison.
  \item Coordinate conversion, if any, is completed before KDS semantics are applied.  For
        example, if external data are read as double precision and then converted to
        binary32, the binary32 boxes are the objects sampled by KDS.
\end{enumerate}

\subsection{Index and counter widths}

If node identifiers, point-array positions, and payloads are stored as 32-bit unsigned
integers, then $n_S<2^{32}$ is required.  If left-side indices or output positions are also
stored in 32-bit fields, then $n_R<2^{32}$ and the per-call sample size must satisfy
$t<2^{32}$.  Larger inputs or outputs require 64-bit fields or chunked processing.

The degree array $w_i$ and total join cardinality $W$ are integer counts.  They should be
stored in a type large enough to represent all possible values, and summation must be
overflow-checked.  Since $W\le n_Rn_S$, an implementation can test this product using a
wider integer type before committing to a fixed-width accumulator.  Silent wraparound
invalidates the sampling distribution.

The integer alias construction may require products of the form $k\omega_h$.  These products
must also be overflow-checked.  If they do not fit the available machine integer type, the
implementation should use a wider type, arbitrary precision, or the prefix-sum weighted
index.

\subsection{Floating-point comparisons}

All KD-tree predicates are closed-interval comparisons on finite point coordinates and query
bounds that may include infinite sentinels.  Because input point coordinates are finite, the
usual comparisons with $-\infty$ and $+\infty$ behave as expected.  NaNs are excluded because
comparisons involving NaN would not define a total ordered predicate.

Bounding boxes are computed from the stored lifted point coordinates by coordinate-wise
minimum and maximum.  They must not be widened or rounded through a different coordinate type
in a way that changes containment or disjointness decisions.  If a wider type is used for
internal comparisons, values should still represent the exact finite coordinates and the
query bounds induced by the chosen input type.

\subsection{Output validation and debugging}

The output phase does not need to recheck geometric intersection for every sampled pair.
Correctness follows from lifting equivalence, exact range counting, item coverage, and exact
sampling.  Debug builds may nevertheless re-evaluate the original strict-overlap predicate
on sampled outputs to detect implementation mistakes in tree construction, query generation,
or payload handling.

\subsection{Chunked sampling}

For very large requested sample sizes, the output can be generated in chunks.  A chunked call
uses the same KDS state and continues the random stream across chunks.  Under the exact
random-source model, generating chunks of sizes $c_1,\ldots,c_q$ is distributionally identical
to one call of size $\sum_h c_h$, because every output position is assigned fresh independent
random choices from the same sampling procedure.

\section{Reproducible Evaluation Protocol}

This section specifies a complete synthetic benchmark protocol.  It is not used by the
correctness proof, but it fixes the data-generation and reporting conventions needed for
comparable experiments.

\subsection{Meaning of the density parameter}

For a balanced synthetic dataset with
\[
  |\Rrel|=|\Srel|=N,
\]
the parameter $\alpha_{\mathrm{out}}$ denotes the target expected number of right-side matches
per left-side box.  Equivalently, the target expected join cardinality is
\[
  \mathbb E[|\Jset|]=N\alpha_{\mathrm{out}},
\]
and the target pairwise intersection probability is
\[
  p_{\mathrm{target}}=\frac{\alpha_{\mathrm{out}}}{N}.
\]
The parameter grid should satisfy $0<p_{\mathrm{target}}<1$.

\subsection{Coordinate distribution}

For both relations and every object, sample a center vector
\[
  c=(c_0,\ldots,c_{d-1})
\]
with independent coordinates uniformly distributed over $[0,1]$.  Boxes are not clipped to
$[0,1]^d$; their lower and upper endpoints may lie outside the unit cube.  This choice
preserves a simple translation model and avoids boundary truncation effects.

Given side lengths $\ell_0,\ldots,\ell_{d-1}>0$, the box is
\[
  \prod_{a=0}^{d-1}[c_a-\ell_a/2,\ c_a+\ell_a/2).
\]
The generated endpoints are rounded to the configured coordinate type, normally binary32,
and boxes failing $\lo<\hi$ after rounding are rejected and regenerated.  With the length
ranges used below, this rejection is not expected except in degenerate parameter settings.

\subsection{Data families}

Two families are used.

\paragraph{F0: isotropic boxes.}
All dimensions use the same side length:
\[
  \ell_0=\cdots=\ell_{d-1}=\lambda.
\]
For continuous centers, the one-dimensional probability that two intervals of length
$\lambda\le 1$ strictly overlap is
\[
  2\lambda-\lambda^2.
\]
Therefore the analytic initial scale for the target density is
\[
  \lambda_0 = 1-\sqrt{1-p_{\mathrm{target}}^{1/d}}.
\]

\paragraph{F1: anisotropic equal-volume boxes.}
Fix an anisotropy ratio $\rho>1$.  The default preset \texttt{TH} uses $\rho=16$.  Define
shape multipliers
\[
  g_0=\rho^{(d-1)/d},
  \qquad
  g_a=\rho^{-1/d}\quad\text{for }a=1,\ldots,d-1.
\]
Then $\prod_{a=0}^{d-1}g_a=1$.  Side lengths are
\[
  \ell_a=\lambda g_a.
\]
Thus all F1 boxes have volume $\lambda^d$, equal to the volume of an F0 box with side length
$\lambda$, but one dimension is elongated while the others are compressed.  The initial
scale $\lambda_0$ is found by bisection over the continuous formula
\[
  \prod_{a=0}^{d-1}(2\lambda g_a-(\lambda g_a)^2)
  = p_{\mathrm{target}},
\]
restricted to $0<\lambda g_a\le 1$ for all $a$.

\subsection{Finite-semantics tuning}

To align the continuous formulas with binary32 endpoint rounding and strict-overlap
semantics, the benchmark uses an empirical tuning pass.  Starting from the analytic initial
scale, bisection searches for a scale $\lambda$ whose estimated mean output degree is within
a relative tolerance of the target.  The default settings are
\[
  \texttt{tune\_samples}=200000,
  \qquad
  \texttt{tune\_tol\_rel}=0.02,
  \qquad
  \texttt{tune\_max\_iter}=30.
\]
For each candidate $\lambda$, the tuning pass draws \texttt{tune\_samples} independent pairs
of boxes using the same finite-coordinate generation rules as the benchmark and estimates
\[
  \widehat\alpha=N\widehat p,
\]
where $\widehat p$ is the empirical strict-overlap probability.  The selected $\lambda$ is
recorded with the dataset metadata.

\subsection{Randomness and seeds}

Each dataset records:
\begin{enumerate}[leftmargin=2em]
  \item the data-generation seed;
  \item the tuning seed;
  \item the sampling seed;
  \item the random-number generator family and version;
  \item the coordinate type;
  \item the final side-length vector.
\end{enumerate}
A counter-based generator or a specified implementation of \texttt{mt19937\_64} may be used.
When centers are converted to binary32, they are generated as binary32 values by rounding the
real-valued uniform sample to the nearest representable coordinate.  All benchmark reports
should include enough metadata to regenerate the same input relation and the same sample
sequence.

\subsection{Binary input format}

A compact little-endian binary format can be used:
\begin{center}
\begin{tabular}{lll}
\toprule
Field & Type & Meaning \\
\midrule
magic & uint32 & fixed file identifier \\
version & uint32 & format version \\
dim & uint32 & dimension $d$ \\
nR & uint64 & number of left-side boxes \\
nS & uint64 & number of right-side boxes \\
coord\_type & uint32 & e.g. 1 for binary32 \\
reserved & uint32 & reserved for alignment \\
R boxes & float32 array & $n_R$ records \\
S boxes & float32 array & $n_S$ records \\
\bottomrule
\end{tabular}
\end{center}
Each box record is stored as
\[
  (\lo[0],\ldots,\lo[d-1],\hi[0],\ldots,\hi[d-1]).
\]
The reader validates the header, dimension, coordinate type, file length, and every box
boundary before building KDS.

\subsection{Timing protocol}

A benchmark run reports separate times for:
\begin{enumerate}[leftmargin=2em]
  \item input loading and validation;
  \item right-side lifting;
  \item KD-tree construction;
  \item degree counting and left-side weighted-index construction;
  \item sample generation.
\end{enumerate}
The commonly reported aggregate fields are
\[
  \mathrm{build\_ms},
  \qquad
  \mathrm{count\_ms},
  \qquad
  \mathrm{sample\_ms},
  \qquad
  \mathrm{total\_ms}.
\]
The exact join size is reported as
\[
  \mathrm{exact\_join\_size}=W.
\]
For a nonzero sample size $t$, throughput fields are
\[
  \mathrm{samples\_per\_sec}=\frac{t}{\mathrm{sample\_ms}/1000},
\]
and
\[
  \mathrm{ns\_per\_sample}=\frac{10^6\cdot\mathrm{sample\_ms}}{t}.
\]
Hardware information, compiler version, compiler flags, thread count, weighted-index type,
grouping method, and chunk size are recorded with every run.

\subsection{Reference parameter grid}

The reference grid uses
\[
  d\in\{2,3,4,5\}
\]
and both data families F0 and F1.

\paragraph{Density sweep.}
Fix
\[
  N=1{,}000{,}000,
  \qquad
  t=10{,}000{,}000,
\]
and use
\[
  \alpha_{\mathrm{out}}\in
  \{0.1,0.3,1,3,10,30,100,300,1000\}.
\]

\paragraph{Input-size sweep.}
Fix
\[
  \alpha_{\mathrm{out}}=100,
  \qquad
  t=10{,}000{,}000,
\]
and use
\[
  N\in
  \{100{,}000,200{,}000,500{,}000,1{,}000{,}000,2{,}000{,}000,5{,}000{,}000\}.
\]

\paragraph{Output-size sweep.}
Fix
\[
  N=1{,}000{,}000,
  \qquad
  \alpha_{\mathrm{out}}=100,
\]
and use
\[
  t\in
  \{1{,}000{,}000,3{,}000{,}000,10{,}000{,}000,30{,}000{,}000,
    100{,}000{,}000,300{,}000{,}000,1{,}000{,}000{,}000\}.
\]
The default chunk size for sample generation is
\[
  \texttt{chunk\_size}=2{,}000{,}000.
\]

\section{Explicit Lifting Examples}

\subsection{Two dimensions}

For $s=[s_x^{\lo},s_x^{\hi})\times[s_y^{\lo},s_y^{\hi})$,
\[
  p(s)=\bigl(s_x^{\lo},s_y^{\lo},s_x^{\hi},s_y^{\hi}\bigr).
\]
For query box $r$,
\[
  q^{\lo}=\bigl(-\infty,-\infty,\nextUp(r_x^{\lo}),\nextUp(r_y^{\lo})\bigr),
\]
\[
  q^{\hi}=\bigl(\nextDown(r_x^{\hi}),\nextDown(r_y^{\hi}),+\infty,+\infty\bigr).
\]

\subsection{Three dimensions}

For a three-dimensional box $s$,
\[
  p(s)=\bigl(s_x^{\lo},s_y^{\lo},s_z^{\lo},s_x^{\hi},s_y^{\hi},s_z^{\hi}\bigr).
\]
For query box $r$,
\[
  q^{\lo}=\bigl(-\infty,-\infty,-\infty,
    \nextUp(r_x^{\lo}),\nextUp(r_y^{\lo}),\nextUp(r_z^{\lo})\bigr),
\]
\[
  q^{\hi}=\bigl(\nextDown(r_x^{\hi}),\nextDown(r_y^{\hi}),\nextDown(r_z^{\hi}),
    +\infty,+\infty,+\infty\bigr).
\]

\subsection{Four dimensions}

For a four-dimensional box $s$,
\[
  p(s)=\bigl(s_0^{\lo},s_1^{\lo},s_2^{\lo},s_3^{\lo},
              s_0^{\hi},s_1^{\hi},s_2^{\hi},s_3^{\hi}\bigr).
\]
For query box $r$,
\[
  q^{\lo}=\bigl(-\infty,-\infty,-\infty,-\infty,
    \nextUp(r_0^{\lo}),\nextUp(r_1^{\lo}),\nextUp(r_2^{\lo}),\nextUp(r_3^{\lo})\bigr),
\]
\[
  q^{\hi}=\bigl(\nextDown(r_0^{\hi}),\nextDown(r_1^{\hi}),
    \nextDown(r_2^{\hi}),\nextDown(r_3^{\hi}),
    +\infty,+\infty,+\infty,+\infty\bigr).
\]

\subsection{Five dimensions}

For a five-dimensional box $s$,
\[
\begin{aligned}
  p(s)=(&s_0^{\lo},s_1^{\lo},s_2^{\lo},s_3^{\lo},s_4^{\lo},\\
        &s_0^{\hi},s_1^{\hi},s_2^{\hi},s_3^{\hi},s_4^{\hi}).
\end{aligned}
\]
For query box $r$,
\[
\begin{aligned}
  q^{\lo}=(&-\infty,-\infty,-\infty,-\infty,-\infty,\\
            &\nextUp(r_0^{\lo}),\nextUp(r_1^{\lo}),\nextUp(r_2^{\lo}),\\
            &\nextUp(r_3^{\lo}),\nextUp(r_4^{\lo})),
\end{aligned}
\]
\[
\begin{aligned}
  q^{\hi}=(&\nextDown(r_0^{\hi}),\nextDown(r_1^{\hi}),\nextDown(r_2^{\hi}),\\
            &\nextDown(r_3^{\hi}),\nextDown(r_4^{\hi}),\\
            &+\infty,+\infty,+\infty,+\infty,+\infty).
\end{aligned}
\]

\section{Conclusion}

KDS reduces exact uniform sampling from a half-open box-intersection join to exact weighted
sampling over left-side degrees and exact conditional sampling over KD-tree range-query
results.  The lift from boxes to $2d$-dimensional points converts strict overlap into a
closed orthogonal range query with exact floating-point boundary semantics.  Range counting
computes the left-side degrees and the total join cardinality.  The item decomposition of a
selected range query gives a disjoint weighted representation of all matching right-side
boxes, allowing each conditional right-side draw to be uniform without materializing the
entire join result.  With exact integer random primitives and overflow-checked count
arithmetic, the output sequence is strictly uniform, independent, and with replacement over
$\Jset$.

\end{document}
